---
title: "Difficulty: Data Exploration and Diagnostics (Likert ratings; treated as approximately continuous)"
author: "CongLiu"
params:
  data_path: "Experimental_Raw_Data.csv"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    word_document: default
    pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  dpi = 300,
  out.width = "90%"
)
options(width = 120)
set.seed(2025)
```

```{r data_paths, echo=FALSE}
# -------------------------------------------------------------------
# Data path helper
# - By default, looks for the data file outside the current analysis folder.
# - Edit `candidate_dirs` if your project uses a different directory layout.
# -------------------------------------------------------------------
locate_data_file <- function(fname){
  if (is.null(fname) || !nzchar(fname)) stop("`data_path` is empty.")
  # 1) Use as-is if it already exists (absolute or relative to current working dir)
  if (file.exists(fname)) return(fname)

  candidate_dirs <- c(
    "..",
    "../data",
    "../datasets",
    "../raw",
    "../input",
    "../../data",
    "../../datasets"
  )

  tried <- c(file.path(getwd(), fname))
  for (d in candidate_dirs){
    f <- file.path(d, fname)
    tried <- c(tried, file.path(getwd(), f))
    if (file.exists(f)) return(f)
  }

  stop(
    "Data file not found: ", fname, "\n",
    "Tried the following locations (relative to the current working directory):\n  - ",
    paste(tried, collapse = "\n  - ")
  )
}
```




# 0. Packages and helper functions

```{r}
need <- c("tidyverse","janitor","readr","dplyr","stringr","forcats","ggplot2","ggpubr",
          "lme4","emmeans","broom.mixed","car","moments","patchwork","knitr","kableExtra","ordinal")
to_install <- setdiff(need, rownames(installed.packages()))
if(length(to_install)) install.packages(to_install, repos = "https://cloud.r-project.org")

suppressPackageStartupMessages({
  library(tidyverse); library(janitor); library(readr); library(dplyr); library(stringr); library(forcats)
  library(ggplot2); library(ggpubr); library(lme4); library(emmeans); library(broom.mixed)
  library(car); library(moments); library(patchwork); library(knitr); library(kableExtra); library(ordinal)
})

coalesce_col <- function(df, patterns, new){
  hit <- names(df)[stringr::str_detect(names(df), stringr::regex(paste0("(", paste(patterns, collapse="|"), ")"), ignore_case=TRUE))]
  if(length(hit) == 0) return(df)
  df %>% dplyr::rename(!!new := dplyr::all_of(hit[1]))
}
```

# 1. Data import and harmonization

```{r}
data_file <- locate_data_file(params$data_path)
raw <- readr::read_csv(data_file, show_col_types = FALSE) %>% janitor::clean_names()

df <- raw %>%
  coalesce_col(c("^participant$","^subj","^pid","^ppid"), "participant") %>%
  coalesce_col(c("^group$","condition","interact","arm"), "group") %>%
  coalesce_col(c("^complexity$","path_complex","task_complexity","simple|complex"), "complexity") %>%
  coalesce_col(c("^difficulty$","^diff_rating$","difficulty_score","diff"), "difficulty") 

stopifnot(all(c("participant","group","complexity","difficulty") %in% names(df)))

df <- df %>%
  mutate(
    participant = factor(participant),
    group = as.factor(group),
    group = ifelse(grepl("2D", group, ignore.case = TRUE), "2D+3D", "3D-only"),
    group = factor(group, levels = c("3D-only","2D+3D")),
    complexity = as.factor(complexity),
    complexity = ifelse(grepl("complex", complexity, ignore.case = TRUE), "complex", "simple"),
    complexity = factor(complexity, levels = c("simple","complex")),
    cell = interaction(group, complexity, sep = "·"),
    difficulty = as.numeric(difficulty)
  )

summary_table <- tibble(
  n_rows = nrow(df),
  n_participants = n_distinct(df$participant)
)

summary_table
```

```{r}
# Frequency distribution (Likert)
tab_diff <- df %>% count(difficulty) %>% mutate(pct = n/sum(n)*100)
kable(tab_diff, digits = 2, caption = "Difficulty rating distribution (overall)") %>% kable_styling(full_width = FALSE)

# Means by condition
by_cell <- df %>% group_by(group, complexity) %>%
  summarise(
    n = n(),
    mean = mean(difficulty, na.rm=TRUE),
    sd = sd(difficulty, na.rm=TRUE)
  )

kable(by_cell, digits = 2, caption = "Descriptive statistics by group × complexity") %>% kable_styling(full_width = FALSE)
```

# 2. Distributional checks (raw Difficulty scale)

```{r fig.height=4.5, fig.width=7}
p1 <- ggplot(df, aes(x=difficulty)) +
  geom_histogram(binwidth = 1, center = 3, boundary = 0.5, closed = "right") +
  scale_x_continuous(breaks = seq(floor(min(df$difficulty, na.rm=TRUE)), ceiling(max(df$difficulty, na.rm=TRUE)), by=1)) +
  theme_minimal() + ggtitle("Histogram of Difficulty ratings (1–5 Likert)")

p2 <- ggplot(df, aes(x=as.factor(difficulty))) +
  geom_bar() + theme_minimal() +
  labs(x = "Difficulty", y = "Count", title = "Bar plot of Difficulty categories") 

p1 / p2
```

```{r}
sk <- moments::skewness(df$difficulty, na.rm=TRUE)
ku <- moments::kurtosis(df$difficulty, na.rm=TRUE) - 3

tibble(
  scale = "raw Difficulty",
  skewness = sk,
  kurtosis_excess = ku
) %>% kable(digits=3, caption="Skewness and kurtosis of Difficulty") %>% kable_styling(full_width=FALSE)
```

# 3. Basic diagnostics for a Gaussian LMM (random intercept only)

```{r}
m_gauss <- lme4::lmer(difficulty ~ group*complexity + (1|participant), data=df, REML=FALSE)
summary(m_gauss)
```

```{r}
diag_df <- broom.mixed::augment(m_gauss) %>%
  mutate(std_resid = .resid / sigma(m_gauss))
```

```{r fig.height=5, fig.width=7}
ggplot(diag_df, aes(.fitted, std_resid)) +
  geom_hline(yintercept = 0, linetype="dashed") +
  geom_point(alpha=.5) +
  theme_minimal() +
  labs(x="Fitted", y="Standardized residuals", title="Difficulty: Gaussian LMM residuals vs fitted values") 
```

```{r fig.height=4.5, fig.width=7}
ggqqplot(diag_df$std_resid, title = "Difficulty: Q–Q plot of standardized residuals (Gaussian LMM)") 
```

```{r}
sw <- shapiro.test(diag_df$std_resid)
tibble(
  model = "Gaussian LMM (Difficulty)",
  W = unname(sw$statistic),
  p_value = sw$p.value
) %>% kable(digits=4, caption="Shapiro–Wilk test for normality (standardized residuals)") %>% kable_styling(full_width=FALSE)
```

# 4. Heteroscedasticity checks (residual variance by cell; Levene test)

```{r}
res_df <- diag_df %>%
  mutate(cell = interaction(group, complexity, sep=":"))

lev <- car::leveneTest(std_resid ~ cell, data = res_df, center = median)
lev
```

```{r}
var_tbl <- res_df %>%
  group_by(cell) %>%
  summarise(resid_var = var(std_resid, na.rm=TRUE), .groups="drop") 

kable(var_tbl, digits=4, caption="Difficulty: residual variance by condition cell") %>% 
  kable_styling(full_width=FALSE)
```

# 5. Notes for subsequent model selection

- In the present dataset, Difficulty is approximately symmetric, with no strong floor/ceiling effects.  
- Gaussian LMM residuals show no severe departures; heteroscedasticity appears modest.  
- Therefore, in the subsequent model comparison we will:  
    - treat Difficulty as an approximately continuous outcome and fit Gaussian LMMs (random intercept vs. random slope);  
    - additionally fit an ordinal mixed model (CLMM) as a sensitivity analysis.

