---
title: "Difficulty: Model Selection (Gaussian LMM vs. CLMM)"
author: "CongLiu"
params:
  data_path: "Experimental_Raw_Data.csv"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  dpi = 300,
  out.width = "90%"
)
options(width = 120)
set.seed(2025)
```

```{r data_paths, echo=FALSE}
# -------------------------------------------------------------------
# Data path helper
# - By default, looks for the data file outside the current analysis folder.
# - Edit `candidate_dirs` if your project uses a different directory layout.
# -------------------------------------------------------------------
locate_data_file <- function(fname){
  if (is.null(fname) || !nzchar(fname)) stop("`data_path` is empty.")
  # 1) Use as-is if it already exists (absolute or relative to current working dir)
  if (file.exists(fname)) return(fname)

  candidate_dirs <- c(
    "..",
    "../data",
    "../datasets",
    "../raw",
    "../input",
    "../../data",
    "../../datasets"
  )

  tried <- c(file.path(getwd(), fname))
  for (d in candidate_dirs){
    f <- file.path(d, fname)
    tried <- c(tried, file.path(getwd(), f))
    if (file.exists(f)) return(f)
  }

  stop(
    "Data file not found: ", fname, "\n",
    "Tried the following locations (relative to the current working directory):\n  - ",
    paste(tried, collapse = "\n  - ")
  )
}
```


# 0. Packages and data

```{r}
need <- c("tidyverse","janitor","lme4","emmeans","broom.mixed","performance","ordinal","knitr","kableExtra")
to_install <- setdiff(need, rownames(installed.packages()))
if(length(to_install)) install.packages(to_install, repos = "https://cloud.r-project.org")

suppressPackageStartupMessages({
  library(tidyverse); library(janitor); library(lme4); library(emmeans); library(broom.mixed)
  library(performance); library(ordinal); library(knitr); library(kableExtra)
})

data_file <- locate_data_file(params$data_path)
df <- readr::read_csv(data_file, show_col_types = FALSE) %>% janitor::clean_names() %>%
  dplyr::rename_with(~"participant", matches("^participant$|^subj|^pid|^ppid", ignore.case=TRUE)) %>%
  dplyr::rename_with(~"group", matches("^group$|condition|interact|arm", ignore.case=TRUE)) %>%
  dplyr::rename_with(~"complexity", matches("^complexity$|path_complex|task_complexity|simple|complex", ignore.case=TRUE)) %>%
  dplyr::rename_with(~"difficulty", matches("^difficulty$|^diff_rating$|difficulty_score|^diff$", ignore.case=TRUE)) %>%
  dplyr::mutate(
    participant = factor(participant),
    group = ifelse(grepl("2D", group, ignore.case=TRUE), "2D+3D", "3D-only"),
    group = factor(group, levels=c("3D-only","2D+3D")),
    complexity = ifelse(grepl("complex", complexity, ignore.case=TRUE), "complex", "simple"),
    complexity = factor(complexity, levels=c("simple","complex")),
    cell = interaction(group, complexity, sep=":"),
    difficulty = as.numeric(difficulty),
    difficulty_ord = ordered(difficulty)  # for CLMM
  )

summary(df$difficulty)
```

# 1. Candidate models

```{r}
# M0: Gaussian LMM, random intercept (primary candidate)
M0 <- lme4::lmer(difficulty ~ group*complexity + (1|participant), data=df, REML=FALSE)

# M1: Gaussian LMM, random intercept + random slope for complexity
M1 <- try(lme4::lmer(difficulty ~ group*complexity + (1 + complexity|participant), data=df, REML=FALSE), silent=TRUE)

# M2: CLMM (ordinal mixed model), random intercept (participant)
M2 <- ordinal::clmm(difficulty_ord ~ group*complexity + (1|participant), data=df, link = "logit", Hess = TRUE)
```

# 2. Information criteria and singularity checks

```{r}
get_info_lmer <- function(m, label){
  if(inherits(m,"try-error")) return(tibble(Model=label, Type="Gaussian LMM", logLik=NA, AIC=NA, BIC=NA, Singular=NA))
  tibble(
    Model = label,
    Type = "Gaussian LMM",
    logLik = as.numeric(logLik(m)),
    AIC = AIC(m),
    BIC = BIC(m),
    Singular = performance::check_singularity(m)  # returns TRUE/FALSE
  )
}

# Other functions and code unchanged
get_info_clmm <- function(m, label){
  tibble(
    Model = label,
    Type = "CLMM (ordinal)",
    logLik = as.numeric(logLik(m)),
    AIC = AIC(m),
    BIC = BIC(m),
    Singular = NA
  )
}

# Rebuild the summary table (subsequent code unchanged)
tab <- dplyr::bind_rows(
  get_info_lmer(M0, "M0: Gaussian LMM (RI)"),
  get_info_lmer(M1, "M1: Gaussian LMM (RI+RS)"),
  get_info_clmm(M2, "M2: CLMM (RI)")
) %>% arrange(AIC)

kable(tab, digits=2, caption="Candidate models for Difficulty: logLik/AIC/BIC and singularity") %>% 
  kable_styling(full_width=FALSE)
```{r DIFF_diag, fig.width=10, fig.height=12, out.width="100%", dpi=300}
  performance::check_model(M0)
```



# 3. Rationale and selection principles (LMM vs. CLMM)

- **Gaussian LMM:** treats Difficulty as approximately continuous; given the near-symmetric distribution and acceptable residual/variance diagnostics, this is the primary analysis choice.  
- **Random slope (M1):** if `Singular = TRUE` or AIC/BIC do not improve, between-participant variability in the simple–complex contrast is limited; a random-intercept model is sufficient.  
- **CLMM (M2):** respects the ordinal nature of Likert responses and is used for **sensitivity analysis**. Information criteria can be inspected, but the key check is whether effect directions and inferential conclusions align with the LMM.  

In the subsequent reporting document (03):  
- the default primary model is **M0 Gaussian LMM (random intercept)**;  
- **M2 CLMM** is additionally fitted and reported as sensitivity analysis.

