---
title: "Head Movement: Data Exploration and Diagnostics (Raw vs Log Scale)"
params:
  data_path: "Experimental_Raw_Data.csv"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  dpi = 300,
  out.width = "90%",
  collapse = TRUE,
  comment = "#>"
)
options(width = 120)
set.seed(2025)

# ------------------------------------------------------------------------------
# Publication-ready data I/O helper
#   - By default, the analysis looks for the data file *outside* this folder,
#     e.g., in ../data/, ../datasets/, or the parent directory.
#   - You can always override `params$data_path` with an absolute/relative path.
# ------------------------------------------------------------------------------
resolve_data_path <- function(
  x,
  candidate_dirs = c("../data", "../datasets", "../dataset", "../raw", "../input", "..")
) {
  if (!is.null(x) && nzchar(x) && file.exists(x)) return(x)

  fname <- basename(x)
  candidates <- unique(c(x, file.path(candidate_dirs, fname)))

  for (p in candidates) {
    if (!is.null(p) && nzchar(p) && file.exists(p)) return(p)
  }

  stop(
    "Data file not found.
",
    "Provided: ", x, "
",
    "Checked paths:
- ", paste(candidates, collapse = "
- ")
  )
}
```


# 0. Packages and helper functions

```{r}
need <- c("tidyverse","janitor","readr","dplyr","stringr","forcats","ggplot2","ggpubr",
          "lme4","nlme","emmeans","broom.mixed","car","moments","patchwork","knitr","kableExtra")
to_install <- setdiff(need, rownames(installed.packages()))
if(length(to_install)) install.packages(to_install, repos = "https://cloud.r-project.org")

suppressPackageStartupMessages({
  library(tidyverse); library(janitor); library(readr); library(dplyr); library(stringr); library(forcats)
  library(ggplot2); library(ggpubr); library(lme4); library(nlme); library(emmeans); library(broom.mixed)
  library(car); library(moments); library(patchwork); library(knitr); library(kableExtra)
})

coalesce_col <- function(df, patterns, new){
  hit <- names(df)[str_detect(names(df), regex(paste0("(", paste(patterns, collapse="|"), ")"), ignore_case=TRUE))]
  if(length(hit) == 0) return(df)
  df %>% rename(!!new := all_of(hit[1]))
}
```

# 1. Data import and standardisation

```{r}
# Resolve data file location
data_path <- resolve_data_path(params$data_path)
raw <- readr::read_csv(data_path, show_col_types = FALSE) %>% janitor::clean_names()

# Try to automatically map key columns
df <- raw %>%
  coalesce_col(c("^participant$","^subj","^pid","^ppid"), "participant") %>%
  coalesce_col(c("^group$","condition","interact","arm"), "group") %>%
  coalesce_col(c("^complexity$","path_complex","task_complexity","simple|complex"), "complexity") %>%
  coalesce_col(c("^head_?move","head.?movement","head.*mov","hm", "head_move_dist"), "head_move")

stopifnot(all(c("participant","group","complexity","head_move") %in% names(df)))

df <- df %>%
  mutate(
    participant = factor(participant),
    # Use inline (?i) to make regex case-insensitive (compatibility with older stringr versions)
    group = as.factor(group) %>% fct_relabel(\(x) ifelse(str_detect(x, regex("2D", ignore_case=TRUE)), "2D+3D", "3D-only")) %>%
              fct_relevel("3D-only", "2D+3D"),
    complexity = as.factor(complexity) %>% fct_relabel(\(x) ifelse(str_detect(x, regex("complex", ignore_case=TRUE)), "complex", "simple")) %>%
                 fct_relevel("simple", "complex"),
    cell = interaction(group, complexity, sep = "·")
  ) %>%
  mutate(head_move = as.numeric(head_move))

# Overall summary
overall_summary <- tibble(
  n_rows = nrow(df),
  n_participants = n_distinct(df$participant)
)

# Number of participants per group
group_summary <- df %>% 
  distinct(participant, group) %>% 
  count(group, name = "n_participants") %>%
  rename(Group = group, Participants = n_participants)

# Trial counts by complexity
complexity_summary <- df %>% 
  count(complexity, name = "n_trials") %>%
  rename(Complexity = complexity, trialCount = n_trials)

# Distribution of trial counts per participant (reporting summary statistics)
trials_per_participant_summary <- df %>% 
  count(participant, name = "n_trials") %>% 
  summarize(
    `Mean trials` = mean(n_trials), 
    `Min trials` = min(n_trials),
    `Max trials` = max(n_trials),
    `SD trials` = sd(n_trials)
  )

# Print all summaries
cat("Overall dataset overview:\n")
print(overall_summary)

cat("\nParticipants per group:\n")
print(group_summary)

cat("\nTrial counts by complexity:\n")
print(complexity_summary)

cat("\nTrial count distribution per participant:\n")
print(trials_per_participant_summary)
```

```{r}
# If any non-positive values are present, apply a minimal shift (should not occur in principle)
eps <- 0
if(any(df$head_move <= 0, na.rm=TRUE)){
  eps <- abs(min(df$head_move[df$head_move> -Inf], na.rm=TRUE)) + 1e-6
}
df <- df %>% mutate(
  head_move_shifted = head_move + eps,
  log_head_move = log(head_move_shifted)
)
```

# 2. Distribution & skewness: raw vs log scale

```{r fig.height=4.5, fig.width=7}
p1 <- ggplot(df, aes(x=head_move)) + geom_histogram(bins=30) + theme_minimal() + ggtitle("Head movement (raw scale)")
p2 <- ggplot(df, aes(x=log_head_move)) + geom_histogram(bins=30) + theme_minimal() + ggtitle("log(HeadMovement)")
p1 / p2
```

```{r}
sk_raw <- moments::skewness(df$head_move, na.rm=TRUE)
ku_raw <- moments::kurtosis(df$head_move, na.rm=TRUE) - 3
sk_log <- moments::skewness(df$log_head_move, na.rm=TRUE)
ku_log <- moments::kurtosis(df$log_head_move, na.rm=TRUE) - 3

tibble(
  scale = c("raw","log"),
  skewness = c(sk_raw, sk_log),
  kurtosis_excess = c(ku_raw, ku_log)
) %>% kable(digits=3) %>% kable_styling(full_width=FALSE)
```

# 3. Diagnostics: LMM on raw scale vs log-LMM

```{r}
m_raw <- lme4::lmer(head_move ~ group * complexity + (1|participant), data=df, REML=FALSE)
m_log <- lme4::lmer(log_head_move ~ group * complexity + (1|participant), data=df, REML=FALSE)

diag_df <- bind_rows(
  broom.mixed::augment(m_raw) %>% mutate(model="Gaussian LMM", std_resid = .resid / sigma(m_raw)),
  broom.mixed::augment(m_log) %>% mutate(model="log-LMM", std_resid = .resid / sigma(m_log))
)
```

```{r fig.height=5.5, fig.width=7}
ggplot(diag_df, aes(.fitted, std_resid)) +
  geom_hline(yintercept = 0, linetype="dashed") +
  geom_point(alpha=.5) +
  facet_wrap(~model, scales="free_x") +
  theme_minimal() + labs(x="Fitted", y="Standardised residuals", title="Residuals vs Fitted")
```

```{r fig.height=5, fig.width=7}
qq1 <- ggqqplot(diag_df %>% filter(model=="Gaussian LMM") %>% pull(std_resid), title = "QQ（Gaussian LMM Standardised residuals）")
qq2 <- ggqqplot(diag_df %>% filter(model=="log-LMM") %>% pull(std_resid), title = "QQ（log-LMM Standardised residuals）")
qq1 / qq2
```

```{r}
# Shapiro test (note: overly sensitive in large samples; for reference only)
sw_raw <- shapiro.test(diag_df %>% filter(model=="Gaussian LMM") %>% pull(std_resid))
sw_log <- shapiro.test(diag_df %>% filter(model=="log-LMM") %>% pull(std_resid))

tibble(
  model = c("Gaussian LMM","log-LMM"),
  W = c(unname(sw_raw$statistic), unname(sw_log$statistic)),
  p_value = c(sw_raw$p.value, sw_log$p.value)
) %>% kable(digits=4) %>% kable_styling(full_width=FALSE)
```

# 4. Heteroscedasticity: Levene tests by cell and variance table

```{r}
res_df <- diag_df %>% select(.resid, std_resid, group, complexity, model) %>% mutate(cell = interaction(group, complexity, sep=":"))
lev_raw <- car::leveneTest(std_resid ~ cell, data = res_df %>% filter(model=="Gaussian LMM"), center = median)
lev_log <- car::leveneTest(std_resid ~ cell, data = res_df %>% filter(model=="log-LMM"), center = median)

lev_raw; lev_log
```

```{r}
var_tbl <- res_df %>%
  group_by(model, cell) %>%
  summarize(resid_var = var(std_resid, na.rm=TRUE), .groups="drop") %>%
  pivot_wider(names_from = cell, values_from = resid_var)

kable(var_tbl, digits=4) %>% kable_styling(full_width=FALSE)
```

# 5. Summary

- On the raw scale, **skewness/kurtosis are pronounced**; standardised residuals show **non-normality and strong heteroscedasticity**.  
- On the log scale, residuals are **closer to normal** and heteroscedasticity is **substantially reduced**.  
- Therefore, downstream modelling is better supported by a **log-LMM** or a **Gamma (log-link)** family.

