---
title: "Head Movement: Model Selection (LMM / varIdent / Gamma-GLMM)"
author: "CongLiu"
params:
  data_path: "Experimental_Raw_Data.csv"
  select_by: "AICw"   # options: "AICw" or "Score"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  dpi = 300,
  out.width = "90%",
  collapse = TRUE,
  comment = "#>"
)
options(width = 120)
set.seed(2025)

# ------------------------------------------------------------------------------
# Publication-ready data I/O helper
#   - By default, the analysis looks for the data file *outside* this folder,
#     e.g., in ../data/, ../datasets/, or the parent directory.
#   - You can always override `params$data_path` with an absolute/relative path.
# ------------------------------------------------------------------------------
resolve_data_path <- function(
  x,
  candidate_dirs = c("../data", "../datasets", "../dataset", "../raw", "../input", "..")
) {
  if (!is.null(x) && nzchar(x) && file.exists(x)) return(x)

  fname <- basename(x)
  candidates <- unique(c(x, file.path(candidate_dirs, fname)))

  for (p in candidates) {
    if (!is.null(p) && nzchar(p) && file.exists(p)) return(p)
  }

  stop(
    "Data file not found.
",
    "Provided: ", x, "
",
    "Checked paths:
- ", paste(candidates, collapse = "
- ")
  )
}
```


# Packages and data

```{r}
need <- c("tidyverse","janitor","lme4","nlme","glmmTMB","emmeans","broom.mixed","performance","insight","see","knitr","kableExtra")
to_install <- setdiff(need, rownames(installed.packages()))
if(length(to_install)) install.packages(to_install, repos = "https://cloud.r-project.org")
suppressPackageStartupMessages({
  library(tidyverse); library(janitor); library(lme4); library(nlme); library(glmmTMB)
  library(emmeans); library(broom.mixed); library(performance); library(insight); library(see)
  library(knitr); library(kableExtra)
})

# Resolve data file location
data_path <- resolve_data_path(params$data_path)
df <- readr::read_csv(data_path, show_col_types = FALSE) %>% janitor::clean_names() %>%
  rename_with(~"participant", matches("^participant$|^subj|^pid|^ppid", ignore.case=TRUE)) %>%
  rename_with(~"group", matches("^group$|condition|interact|arm", ignore.case=TRUE)) %>%
  rename_with(~"complexity", matches("^complexity$|path_complex|task_complexity|simple|complex", ignore.case=TRUE)) %>%
  rename_with(~"head_move", matches("^head_?move|head.?movement|head.*mov|^hm$|head_move_dist", ignore.case=TRUE)) %>%
  mutate(
    participant = factor(participant),
    # Use regex(ignore_case=TRUE) for stable behaviour under older stringr versions
    group = factor(ifelse(str_detect(group, regex("2D", ignore_case=TRUE)), "2D+3D", "3D-only"),
                   levels = c("3D-only","2D+3D")),
    complexity = factor(ifelse(str_detect(complexity, regex("complex", ignore_case=TRUE)), "complex", "simple"),
                        levels = c("simple","complex")),
    cell = interaction(group, complexity, sep = ":")
  ) %>%
  mutate(head_move = as.numeric(head_move))

# Ensure strictly positive values (required for Gamma family and log-transform)
eps <- 0
if(any(df$head_move <= 0, na.rm=TRUE)) eps <- abs(min(df$head_move, na.rm=TRUE)) + 1e-6
df <- df %>% mutate(log_head_move = log(head_move + eps))
```

# Fit candidate models

```{r models}
# M0: Gaussian LMM on raw scale (baseline)
M0 <- lme4::lmer(head_move ~ group*complexity + (1|participant), data=df, REML=FALSE)

# M1: log-LMM (random intercept + random slope)
M1 <- try(lme4::lmer(log_head_move ~ group*complexity + (1 + complexity|participant), data=df, REML=FALSE), silent=TRUE)

# M1a: log-LMM (random intercept only)
M1a <- lme4::lmer(log_head_move ~ group*complexity + (1|participant), data=df, REML=FALSE)

# M2: log-LMM + varIdent (cell-specific residual variance)
M2 <- nlme::lme(fixed = log_head_move ~ group*complexity,
                random = ~1|participant, data=df, method="ML",
                weights = nlme::varIdent(form = ~1|cell))

# M3: Gamma GLMM (log link)
M3 <- glmmTMB::glmmTMB(head_move ~ group*complexity + (1|participant),
                       data=df, family=Gamma(link="log"))

# M3b: Gamma GLMM + dispformula (cell-specific dispersion)
M3b <- glmmTMB::glmmTMB(head_move ~ group*complexity + (1|participant),
                        dispformula = ~ cell,
                        data=df, family=Gamma(link="log"))
```

# Core model metrics (AIC weight / composite performance score / R² / RMSE / sigma)

```{r compare}
# Collect successfully fitted models
mods <- list(M0=M0, M1=M1, M1a=M1a, M2=M2, M3=M3, M3b=M3b)
ok <- !vapply(mods, inherits, logical(1), "try-error")
mods <- mods[ok]

# Basic information criteria
aic <- sapply(mods, AIC)
bic <- sapply(mods, BIC)
ll  <- sapply(mods, logLik)

# R² (marginal), RMSE, sigma
get_r2m <- function(m){
  # For mixed models, use Nakagawa R²
  if (inherits(m, c("lmerMod", "lmerModLmerTest", "lme", "glmmTMB"))) {
    r2 <- try(performance::r2_nakagawa(m), silent = TRUE)
    if (inherits(r2, "try-error")) return(NA_real_)
    return(unname(r2$R2_marginal))
  }

  # For models without random effects (e.g., lm/glm), use r2()
  r2 <- try(performance::r2(m), silent = TRUE)
  if (inherits(r2, "try-error")) return(NA_real_)
  if ("R2_marginal" %in% names(r2)) return(unname(r2[["R2_marginal"]]))
  if ("R2" %in% names(r2))        return(unname(r2[["R2"]]))
  NA_real_
}
r2m   <- sapply(mods, get_r2m)

get_rmse <- function(m){
  rm <- try(performance::rmse(m), silent=TRUE)
  if(inherits(rm, "try-error")) return(NA_real_)
  # rmse() may return a vector or a data frame
  if(is.list(rm) && "RMSE" %in% names(rm)) return(as.numeric(rm$RMSE))
  as.numeric(rm)[1]
}
rmse  <- sapply(mods, get_rmse)

get_sigma <- function(m){
  s <- try(performance::sigma(m), silent=TRUE)
  if(inherits(s, "try-error")) return(NA_real_)
  as.numeric(s)[1]
}
sigma <- sapply(mods, get_sigma)

# Singularity (only applicable to lmer models)
sing <- sapply(mods, function(m){
  if(inherits(m, "lmerMod") || inherits(m, "lmerModLmerTest")){
    res <- try(performance::check_singularity(m), silent=TRUE)
    if(inherits(res, "try-error")) return(NA)
    if(is.atomic(res)) return(as.logical(res))
    if(is.data.frame(res) && "is_singular" %in% colnames(res)) return(res$is_singular)
  }
  NA
})

# AIC weights
deltaAIC <- aic - min(aic, na.rm=TRUE)
aw <- exp(-0.5 * deltaAIC)
AICw <- aw / sum(aw, na.rm=TRUE)

# Composite performance score (0–100): min–max normalise each metric and combine with direction-aware weights
# Direction: larger R² is better; smaller AIC/BIC/RMSE/sigma are better
norm_minmax <- function(x, larger_better=TRUE){
  if(all(is.na(x))) return(rep(NA_real_, length(x)))
  xmin <- min(x, na.rm=TRUE); xmax <- max(x, na.rm=TRUE)
  if(isTRUE(all.equal(xmin, xmax))) return(rep(1, length(x)))  # All values identical -> score = 1
  if(larger_better){
    (x - xmin) / (xmax - xmin)
  } else {
    (xmax - x) / (xmax - xmin)
  }
}
w <- c(AIC=0.30, BIC=0.10, R2=0.30, RMSE=0.20, Sigma=0.10)  # adjust if needed
score <- w["AIC"]  * norm_minmax(aic,  larger_better=FALSE) +
         w["BIC"]  * norm_minmax(bic,  larger_better=FALSE) +
         w["R2"]   * norm_minmax(r2m,  larger_better=TRUE)  +
         w["RMSE"] * norm_minmax(rmse, larger_better=FALSE) +
         w["Sigma"]* norm_minmax(sigma,larger_better=FALSE)
ScorePct <- 100 * as.numeric(score)

# Summary table (sorted by AIC weight and score)
tab_metrics <- tibble::tibble(
  Model = names(mods),
  Class = vapply(mods, function(x) class(x)[1], character(1)),
  logLik = as.numeric(ll),
  AIC    = as.numeric(aic),
  BIC    = as.numeric(bic),
  AICw   = as.numeric(AICw),
  R2_marginal = as.numeric(r2m),
  RMSE   = as.numeric(rmse),
  Sigma  = as.numeric(sigma),
  Singular = as.logical(sing),
  `Performance-Score(%)` = round(ScorePct, 2)
) %>%
  arrange(desc(AICw), desc(`Performance-Score(%)`))

kable(tab_metrics, digits=3, caption="Key metrics for candidate models (incl. AIC weights and composite score)") %>% kable_styling(full_width=FALSE)
```

## `performance::compare_performance` (official comparison table with ranks)

```{r perf_compare}
comp <- performance::compare_performance(mods, rank = TRUE)

# (1) Keep model names as a column (collision-safe)
if (!"Model" %in% names(comp)) {
  comp <- tibble::rownames_to_column(comp, "Model")
}

# (2) If `Rank`/`rank` is absent, generate it
if (!any(c("Rank","rank") %in% names(comp))) {
  comp <- comp %>% 
    mutate(Rank = dplyr::row_number())          # Use current row order as Rank
}

# (3) Harmonise `rank`/`Rank` naming
if ("rank" %in% names(comp) && !"Rank" %in% names(comp)) {
  comp <- dplyr::rename(comp, Rank = rank)
}

comp2 <- comp %>% 
  left_join(tab_metrics %>% 
              dplyr::select(Model, AICw, `Performance-Score(%)`),
            by = "Model") %>% 
  arrange(Rank)                                  # Ensure Rank exists

kable(comp2, digits = 3,
      caption = "compare_performance + AIC weights + composite score") %>% 
  kable_styling(full_width = FALSE)


```

# Automated “best model” selection with rationale

```{r pick_best, results='asis'}
best_by_AICw  <- tab_metrics %>% slice_max(AICw, n=1, with_ties = FALSE) %>% pull(Model)
best_by_Score <- tab_metrics %>% slice_max(`Performance-Score(%)`, n=1, with_ties = FALSE) %>% pull(Model)

best_choice <- if (tolower(params$select_by) == "score") best_by_Score else best_by_AICw
best_row <- tab_metrics %>% filter(Model == best_choice)

cat("**Selection criterion**: ", params$select_by, "<br/>",
    "**Best by AIC weight**: ", best_by_AICw, "<br/>",
    "**Best by composite score**: ", best_by_Score, "<br/>",
    "**Final model used for downstream reporting**: ", best_choice, "<br/><br/>", sep="")

kable(best_row, digits=3, caption = "Key metrics of the selected model") %>% kable_styling(full_width=FALSE)
```

# Diagnostics (quick check of residuals etc. for the selected model)

```{r diag}
best_mod <- mods[[ best_choice ]]
# One-shot diagnostic plots (normality, heteroscedasticity, influence, fit)
performance::check_model(best_mod)
```

# Add-on: information criteria table (for alignment with earlier versions)

```{r aic_table}
AICtab <- tab_metrics %>% select(Model, AIC, BIC, AICw, R2_marginal, RMSE, Sigma, Singular, `Performance-Score(%)`)
kable(AICtab, digits=3, caption="Summary of information criteria and core metrics (incl. weights and score)") %>% kable_styling(full_width=FALSE)
```

# Summary and recommendations

- **AIC weight (AICw)** reflects the relative probability that a model is best among the candidates; values closer to 1 indicate stronger support.  
- **Performance score** is a weighted average (0–100) over normalised metrics (AIC/BIC/R²/RMSE/sigma). If you need the same weighting scheme as in the completion-time analysis, adjust `w` in the code.  
- **Diagnostic plots** are recommended for inclusion in the report (or at least retained in the Supplement). Consider noting that the head-movement model choice was motivated by distributional/variance features and validated via AIC weights and the composite score.

