---
title: "Head Movement: Main Analysis and Reporting (EMMs + Post-hoc Comparisons)"
author: "Supplementary Material"
params:
  data_path: "Experimental_Raw_Data.csv"
  use_varIdent_as_main: true
  do_cluster_boot: false
  boot_B: 500
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
  word_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE, fig.align = 'center', fig.width = 7, fig.height = 5, dpi = 300)
set.seed(2025)
```

# 0. Packages and data

```{r}
need <- c("tidyverse","janitor","lme4","nlme","glmmTMB","emmeans","broom.mixed","modelr","knitr","kableExtra","ggplot2","gghalves","pBrackets","performance")
to_install <- setdiff(need, rownames(installed.packages()))
if(length(to_install)) install.packages(to_install, repos = "https://cloud.r-project.org")
suppressPackageStartupMessages({library(tidyverse);library(janitor);library(lme4);library(nlme);library(glmmTMB);library(emmeans);library(broom.mixed);library(modelr);library(knitr);library(kableExtra);library(ggplot2);library(gghalves);library(pBrackets);library(performance)})

# ---- Data input ------------------------------------------------------------
# The raw data file is expected to live *outside* the analysis folder.
# By default, the script searches common sibling/parent locations.
resolve_data_path <- function(path, candidate_dirs = c('..', '../data', '../datasets')) {
  if (file.exists(path)) return(normalizePath(path, mustWork = FALSE))
  fname <- basename(path)
  candidates <- file.path(candidate_dirs, fname)
  hit <- candidates[file.exists(candidates)][1]
  if (!is.na(hit)) return(normalizePath(hit, mustWork = FALSE))
  stop(
    paste0('Data file not found. Tried:\n- ', paste(c(path, candidates), collapse='\n- '),
           '\n\nTip: set params$data_path to an absolute path, or place the file in one of the candidate folders.')
  )
}
data_file <- resolve_data_path(params$data_path)

df <- readr::read_csv(data_file, show_col_types = FALSE) %>% janitor::clean_names() %>%
  rename_with(~"participant", matches("^participant$|^subj|^pid|^ppid", ignore.case=TRUE)) %>%
  rename_with(~"group", matches("^group$|condition|interact|arm", ignore.case=TRUE)) %>%
  rename_with(~"complexity", matches("^complexity$|path_complex|task_complexity|simple|complex", ignore.case=TRUE)) %>%
  rename_with(~"head_move", matches("^head_?move|head.?movement|head.*mov|^hm$|head_move_dist", ignore.case=TRUE)) %>%
  mutate(
    participant = factor(participant),
    # Use base::grepl(...) instead of stringr::str_detect(..., ignore_case=TRUE) to avoid compatibility issues.
    group = factor(ifelse(grepl("2D", group, ignore.case = TRUE), "2D+3D", "3D-only"), levels = c("3D-only","2D+3D")),
    complexity = factor(ifelse(grepl("complex", complexity, ignore.case = TRUE), "complex", "simple"), levels = c("simple","complex")),
    cell = interaction(group, complexity, sep = ":")
  ) %>%
  mutate(head_move = as.numeric(head_move))

eps <- 0
if(any(df$head_move <= 0, na.rm=TRUE)) eps <- abs(min(df$head_move, na.rm=TRUE)) + 1e-6
df <- df %>% mutate(log_head_move = log(head_move + eps))
```

# 1. Primary model specification (default: M2 log-LMM + varIdent)

- Following the model-selection step (Script 02), the default primary model is **`nlme::lme` with `varIdent` (M2)**.  
- If the `varIdent` model fails to converge or is unstable, the script automatically falls back to a **log-LMM with random intercepts (RI)**.

```{r}
main_tag <- NULL
main_fit <- NULL

if(isTRUE(params$use_varIdent_as_main)){
  try({
    main_fit <- nlme::lme(fixed = log_head_move ~ group*complexity,
                          random = ~1|participant, data=df, method="REML",
                          weights = nlme::varIdent(form = ~1|cell))
    main_tag <- "log-LMM + varIdent (M2)"
  }, silent = TRUE)
}

if(is.null(main_fit)){
  message("Primary varIdent model failed or disabled; falling back to log-LMM (random intercept).")
  main_fit <- lme4::lmer(log_head_move ~ group*complexity + (1|participant), data=df, REML=TRUE)
  main_tag <- "log-LMM (RI)"
}

summary(main_fit)



```

```{r tab2_fixef, results='asis', message=FALSE, warning=FALSE}
fe_table <- function(fit, reference="3D-only × simple", digits=3){
  if (!requireNamespace("parameters", quietly = TRUE)) install.packages("parameters", repos="https://cloud.r-project.org")
  suppressPackageStartupMessages({ library(parameters); library(dplyr) })
  tab <- parameters::model_parameters(fit, ci_method = "wald", verbose = FALSE)
  stat_col <- if("t" %in% names(tab)) "t" else if("z" %in% names(tab)) "z" else if("Statistic" %in% names(tab)) "Statistic" else names(tab)[1]
  out <- tab %>% dplyr::transmute(
    Term = Parameter, `β` = Coefficient, SE = SE,
    Stat = .data[[stat_col]], p = p,
    `95% CI` = sprintf("[%.3f, %.3f]", CI_low, CI_high)
  )
  kableExtra::kable(out, digits = digits,
    caption = paste0("Fixed-effects summary (", main_tag, ") — reference: ", reference)) %>% 
    kableExtra::kable_styling(full_width = FALSE)
}
fe_table(main_fit)
```



## Head movement — Trial-level residual variability (SD ± 95% CI; log scale)

```{r hm_residual_sd_raw, message=FALSE, warning=FALSE, fig.width=7, fig.height=4.5}
# Assumes the primary model object is `main_fit`, and the data include `group` and `complexity`.
suppressPackageStartupMessages({ library(dplyr); library(ggplot2); library(forcats) })

dat <- nlme::getData(main_fit)
dat$cell <- interaction(dat$group, dat$complexity, sep = " × ")

# Residuals on the response scale (for HM this corresponds to the log scale used in the model).
dat$res_raw <- resid(main_fit, type = "response")

sd_by_cell <- dat %>%
  group_by(cell) %>%
  summarise(n = n(), sd = sd(res_raw, na.rm = TRUE), .groups = "drop") %>%
  mutate(
    df = n - 1, alpha = 0.05,
    sd_low  = sqrt(df * sd^2 / qchisq(1 - alpha/2, df)),
    sd_high = sqrt(df * sd^2 / qchisq(alpha/2,   df))
  )

ggplot(sd_by_cell, aes(x = sd, y = fct_rev(cell), xmin = sd_low, xmax = sd_high)) +
  geom_pointrange() +
  labs(x = "Residual SD (log scale)", y = NULL,
       title = "Head movement: Trial-level residual variability by cell (SD ± 95% CI; log scale)") +
  theme_minimal(base_size = 12)
```


## Head movement — Model-based residual SD by cell (from varIdent)
```{r hm_residual_sd_varIdent, message=FALSE, warning=FALSE, fig.width=7, fig.height=4.5}
suppressPackageStartupMessages({ library(nlme); library(dplyr); library(ggplot2); library(forcats) })

# Run this block only when the primary model is an `lme` with `varIdent` enabled.
if (inherits(main_fit, "lme") && !is.null(main_fit$modelStruct$varStruct)) {
  dat <- nlme::getData(main_fit)
  dat$cell <- interaction(dat$group, dat$complexity, sep = " × ")

  # Extract the reference residual SD from the `lme` summary (avoids version-specific sigma methods).
  sigma0 <- summary(main_fit)$sigma  # reference-cell residual SD for the lme model
  # Extract varIdent weights (compatible across nlme versions).
  vw <- coef(main_fit$modelStruct$varStruct, unconstrained = FALSE)  # weights for the other cells (>0)

  # Reference cell label (must match the factor levels in the data).
  ref <- "3D-only × simple"

  cells <- levels(dat$cell)
  w <- setNames(rep(NA_real_, length(cells)), cells)
  w[names(vw)] <- vw
  w[ref] <- 1

  # varIdent relationship: SD_cell = sigma0 / w_cell
  sd_cell <- sigma0 / w
  df_sd <- data.frame(cell = names(sd_cell), sd = as.numeric(sd_cell))

  ggplot(df_sd, aes(x = sd, y = fct_rev(cell))) +
    geom_point(size = 2) +
    labs(x = "Estimated residual SD by varIdent (log scale)",
         y = NULL,
         title = "Head movement: Residual SD by cell inferred from varIdent (larger = more variability)") +
    theme_minimal(base_size = 12)
} else {
  # If the primary model is not `lme` / has no varIdent, emit a note and output a blank plot.
  message("Primary model is not nlme::lme or varIdent is not enabled; skipping varIdent residual-SD plot.")
  # Output a blank plot to avoid errors.
  ggplot() + 
    geom_blank() + 
    labs(title = "varIdent residual-SD plot not generated",
         subtitle = "Reason: primary model is log-LMM (RI) or varIdent is not enabled") +
    theme_minimal()
}
```

## Forest data — standardized effects (Std. β)

```{r forest_std_hm, results='asis', message=FALSE, warning=FALSE}
# Required packages
pkgs <- c("parameters","effectsize","insight","dplyr","knitr")
for(p in pkgs) if(!requireNamespace(p, quietly=TRUE)) install.packages(p, repos="https://cloud.r-project.org")
suppressPackageStartupMessages({ library(parameters); library(effectsize); library(insight); library(dplyr); library(knitr) })

std_table <- function(fit, terms = c("group2D+3D","complexitycomplex","group2D+3D:complexitycomplex"), component=NULL){
  # Try refit-based standardization first
  tab <- NULL
  try({
    if(inherits(fit,"glmmTMB")){
      tab <- effectsize::standardize_parameters(fit, method="refit",
                                                component=if(is.null(component)) "conditional" else component, ci=0.95)
    } else {
      tab <- effectsize::standardize_parameters(fit, method="refit", ci=0.95)
    }
  }, silent = TRUE)

  recode_label <- c("group2D+3D"="Group (2D+3D vs 3D-only)",
                    "complexitycomplex"="Complexity (complex vs simple)",
                    "group2D+3D:complexitycomplex"="Interaction")

  use_refit <- !is.null(tab)
  if(use_refit){
    df <- as.data.frame(tab)
    # Keep only requested terms; if Std_Coefficient is all NA, fall back
    if(all(is.na(df$Std_Coefficient[df$Parameter %in% terms]))){
      use_refit <- FALSE
    } else {
      out <- df[df$Parameter %in% terms, , drop=FALSE]
      out <- out %>%
        dplyr::transmute(Effect = dplyr::recode(Parameter, !!!recode_label),
                         `Std. β` = Std_Coefficient,
                         `CI low`  = CI_low_Std,
                         `CI high` = CI_high_Std)
      return(out)
    }
  }


  raw <- parameters::model_parameters(fit,
          component = if(inherits(fit,"glmmTMB")) "conditional" else NULL,
          ci_method = "wald")
  raw <- as.data.frame(raw)
  raw <- raw[raw$Parameter %in% terms, , drop=FALSE]

  # SD(Y) from data
  mf <- insight::get_data(fit, verbose=FALSE)
  resp <- insight::find_response(fit)
  SDy <- stats::sd(mf[[resp]], na.rm=TRUE)

  # Use *model matrix* so column names match coefficient labels
  mm <- try(insight::get_modelmatrix(fit, component = "conditional"), silent = TRUE)
  if(inherits(mm, "try-error") || is.null(mm)){
    mm <- try(stats::model.matrix(stats::formula(fit), data = mf), silent = TRUE)
  }

  get_sdx <- function(param){
    if(!is.null(mm) && !inherits(mm, "try-error") && param %in% colnames(mm)){
      return(stats::sd(mm[, param], na.rm=TRUE))
    }
    # Last resort: approximate using predictor variance if column can't be found
    varname <- sub(":.*", "", sub("(.*?)(\\w+)$", "\\1\\2", param))
    if(!is.null(mf[[varname]])){
      x <- mf[[varname]]
      ux <- stats::na.omit(unique(x))
      if(length(ux)==2 && all(sort(ux) %in% c(0,1))) {
        p <- mean(x, na.rm=TRUE); return(sqrt(p*(1-p)))
      }
      return(stats::sd(x, na.rm=TRUE))
    }
    return(NA_real_)
  }

  sdx <- vapply(raw$Parameter, get_sdx, numeric(1))
  out <- raw %>%
    dplyr::transmute(Effect = dplyr::recode(Parameter, !!!recode_label),
                     `Std. β` = Coefficient * sdx / SDy,
                     `CI low` = CI_low     * sdx / SDy,
                     `CI high`= CI_high    * sdx / SDy)
  return(out)
}

# --- Choose the model object for this document ---
forest_df <- std_table(main_fit, component=NULL)

knitr::kable(forest_df, digits=3,
             caption = "Forest data — standardized effects (Std. β with 95% CI; baseline = 3D-only × simple)")
```

```

# 2. Estimated marginal means (EMMs) and pairwise comparisons (Holm)

> The model is fit on the **log scale**; reported means are **exponentiated** to yield **geometric means** on the original scale (for interpretability).

```{r}
emm_log <- emmeans::emmeans(main_fit, ~ group*complexity)

# EMMs (geometric means)
emm_tbl <- summary(emm_log, infer=c(TRUE, TRUE)) %>%
  as.data.frame() %>%
  mutate(geom_mean = exp(emmean),
         lower.CL.exp = exp(lower.CL),
         upper.CL.exp = exp(upper.CL)) %>%
  select(group, complexity, emmean, SE, df, lower.CL, upper.CL, geom_mean, lower.CL.exp, upper.CL.exp) %>%
  arrange(group, complexity)

kable(emm_tbl, digits=3, caption = paste0("Estimated marginal means (", main_tag, "): geometric mean and 95% CI (original scale)")) %>% kable_styling(full_width=FALSE)
```

```{r}
# Pairwise comparisons across two factors (Holm adjustment)
cmp_group_within_complex <- pairs(emm_log, by="complexity", adjust="holm")
cmp_complex_within_group <- pairs(emm_log, by="group", adjust="holm")

cmp_group_tbl <- summary(cmp_group_within_complex, infer=c(TRUE, TRUE)) %>%
  as.data.frame() %>%
  mutate(estimate_ratio = exp(estimate),
         lower.CL.exp = exp(lower.CL),
         upper.CL.exp = exp(upper.CL)) %>%
  select(complexity, contrast, estimate, SE, df, p.value, estimate_ratio, lower.CL.exp, upper.CL.exp)

cmp_complex_tbl <- summary(cmp_complex_within_group, infer=c(TRUE, TRUE)) %>%
  as.data.frame() %>%
  mutate(estimate_ratio = exp(estimate),
         lower.CL.exp = exp(lower.CL),
         upper.CL.exp = exp(upper.CL)) %>%
  select(group, contrast, estimate, SE, df, p.value, estimate_ratio, lower.CL.exp, upper.CL.exp)

kable(cmp_group_tbl, digits=3, caption="GMR: 2D+3D vs 3D-only (stratified by complexity)") %>% kable_styling(full_width=FALSE)
kable(cmp_complex_tbl, digits=3, caption="GMR: complex vs simple (stratified by group)") %>% kable_styling(full_width=FALSE)
```

# 3. Visualization: geometric means across the four cells

```{r hm_raincloud, fig.height=7, fig.width=7}
emm_ct_df <- emm_tbl %>%
  dplyr::transmute(
    group, complexity,
    time_sec  = geom_mean,
    lower_sec = lower.CL.exp,
    upper_sec = upper.CL.exp
  )

dat_ct_raw <- df %>%
  dplyr::select(participant, group, complexity, head_move) %>%
  dplyr::rename(time_sec = head_move)

# ---------- Colors ----------
col_blue <- "#3E69B0"  # 3D-only
col_red  <- "#C61B21"  # 2D+3D

# ---------- Data range ----------
y_raw_min <- min(dat_ct_raw$time_sec, emm_ct_df$lower_sec, na.rm = TRUE)
y_raw_max <- max(dat_ct_raw$time_sec, emm_ct_df$upper_sec, na.rm = TRUE)

# Y-axis ticks (omit 0)
y_breaks_all  <- pretty(c(y_raw_min, y_raw_max), n = 5)
y_breaks_used <- y_breaks_all

y_axis_min <- min(y_breaks_used)
y_axis_max <- max(y_breaks_used)
dy         <- y_axis_max - y_axis_min

# Plot limits (y-axis)
y_plot_min <- y_axis_min - 0.30 * dy
y_plot_max <- y_axis_max + 0.05 * dy

# ---------- X-axis numeric positions ----------
x_simple  <- 1
x_complex <- 2
x_rng     <- x_complex - x_simple

x_plot_min <- x_simple - 0.7
x_plot_max <- x_complex + 0.5

# ---------- Custom y-axis ticks (left-facing) ----------
y_axis_x  <- x_simple - 0.35        # vertical axis x
y_tick_x1 <- y_axis_x               # tick start (on axis line)
y_tick_x2 <- y_axis_x - 0.05        # tick end (to the left)

# ---------- Custom x-axis ticks (down-facing) ----------
x_axis_y  <- y_axis_min - 0.08 * dy       # horizontal axis y
x_tick_y1 <- x_axis_y                     # tick start
x_tick_y2 <- x_axis_y - 0.04 * dy         # tick end (downwards)

# =====================================================================
# Main panel
# =====================================================================
p_ct_raincloud <- ggplot() +

  # ===== simple side: left half-violin + points + boxplot =====
  gghalves::geom_half_violin(
    data = dat_ct_raw %>% filter(complexity == "simple", group == "3D-only"),
    aes(x = as.numeric(complexity), y = time_sec,
        fill = group, colour = group),
    side = "l", width = 0.4, trim = FALSE,
    alpha = 0.35, linewidth = 0.9
  ) +
  gghalves::geom_half_violin(
    data = dat_ct_raw %>% filter(complexity == "simple", group == "2D+3D"),
    aes(x = as.numeric(complexity), y = time_sec,
        fill = group, colour = group),
    side = "l", width = 0.4, trim = FALSE,
    alpha = 0.35, linewidth = 0.9
  ) +
  geom_point(
    data = dat_ct_raw %>% filter(complexity == "simple"),
    aes(x = as.numeric(complexity) + 0.10, y = time_sec, colour = group),
    position = position_jitter(width = 0.05, height = 0, seed = 2025),
    alpha = 0.6, shape = 16, size = 2.2
  ) +
  geom_boxplot(
    data = dat_ct_raw %>% filter(complexity == "simple", group == "3D-only"),
    aes(x = as.numeric(complexity) + 0.25, y = time_sec,
        fill = group, colour = group),
    width = 0.08,
    alpha = 0.25, outlier.shape = NA, linewidth = 0.9
  ) +
  geom_boxplot(
    data = dat_ct_raw %>% filter(complexity == "simple", group == "2D+3D"),
    aes(x = as.numeric(complexity) + 0.35, y = time_sec,
        fill = group, colour = group),
    width = 0.08,
    alpha = 0.25, outlier.shape = NA, linewidth = 0.9
  ) +

  # ===== complex side: right half-violin + points + boxplot =====
  gghalves::geom_half_violin(
    data = dat_ct_raw %>% filter(complexity == "complex", group == "3D-only"),
    aes(x = as.numeric(complexity), y = time_sec,
        fill = group, colour = group),
    side = "r", width = 0.4, trim = FALSE,
    alpha = 0.35, linewidth = 0.9
  ) +
  gghalves::geom_half_violin(
    data = dat_ct_raw %>% filter(complexity == "complex", group == "2D+3D"),
    aes(x = as.numeric(complexity), y = time_sec,
        fill = group, colour = group),
    side = "r", width = 0.4, trim = FALSE,
    alpha = 0.35, linewidth = 0.9
  ) +
  geom_point(
    data = dat_ct_raw %>% filter(complexity == "complex"),
    aes(x = as.numeric(complexity) - 0.10, y = time_sec, colour = group),
    position = position_jitter(width = 0.05, height = 0, seed = 2025),
    alpha = 0.6, shape = 16, size = 2.2
  ) +
  geom_boxplot(
    data = dat_ct_raw %>% filter(complexity == "complex", group == "3D-only"),
    aes(x = as.numeric(complexity) - 0.35, y = time_sec,
        fill = group, colour = group),
    width = 0.08,
    alpha = 0.25, outlier.shape = NA, linewidth = 0.9
  ) +
  geom_boxplot(
    data = dat_ct_raw %>% filter(complexity == "complex", group == "2D+3D"),
    aes(x = as.numeric(complexity) - 0.25, y = time_sec,
        fill = group, colour = group),
    width = 0.08,
    alpha = 0.25, outlier.shape = NA, linewidth = 0.9
  ) +

  # ===== EMM =====
  geom_line(
  data = emm_ct_df,
  aes(
    x = as.numeric(complexity) + case_when(
      # Compute the horizontal offsets used for each group×complexity cell (to match the boxplots).
      group == "3D-only" & complexity == "simple" ~ 0.25,  # boxplot offset+0.25
      group == "2D+3D" & complexity == "simple" ~ 0.35,  # boxplot offset+0.35
      group == "3D-only" & complexity == "complex" ~ -0.35, # boxplot offset-0.35
      group == "2D+3D" & complexity == "complex" ~ -0.25  # boxplot offset-0.25
    ),
    y = time_sec,
    group = group,
    colour = group
  ),
  linewidth = 1.1
) +
geom_errorbar(
  data = emm_ct_df,
  aes(
    x = as.numeric(complexity) + case_when(
      group == "3D-only" & complexity == "simple" ~ 0.25,
      group == "2D+3D" & complexity == "simple" ~ 0.35,
      group == "3D-only" & complexity == "complex" ~ -0.35,
      group == "2D+3D" & complexity == "complex" ~ -0.25
    ),
    ymin = lower_sec, ymax = upper_sec,
    group = group, colour = group
  ),
  width = 0, linewidth = 1.1
) +
geom_point(
  data = emm_ct_df,
  aes(
    x = as.numeric(complexity) + case_when(
      group == "3D-only" & complexity == "simple" ~ 0.25,
      group == "2D+3D" & complexity == "simple" ~ 0.35,
      group == "3D-only" & complexity == "complex" ~ -0.35,
      group == "2D+3D" & complexity == "complex" ~ -0.25
    ),
    y = time_sec,
    group = group,
    fill = group,
    colour = group
  ),
  shape = 23, size = 3.0, stroke = 1.0
) +
 # ===== min–max line for all raw data (with end caps) =====
  stat_summary(
    data = dat_ct_raw,
    aes(
      x = as.numeric(complexity) + dplyr::case_when(
        group == "3D-only" & complexity == "simple"  ~  0.25,
        group == "2D+3D"  & complexity == "simple"  ~  0.35,
        group == "3D-only" & complexity == "complex" ~ -0.35,
        group == "2D+3D"  & complexity == "complex" ~ -0.25
      ),
      y     = time_sec,
      colour = group,
      group  = interaction(group, complexity)  # one line per group×complexity
    ),
    fun.min = min,      # lower end: min of raw data
    fun.max = max,      # upper end: max of raw data
    geom    = "errorbar",
    width   = 0.07,     # >0 draws end caps; adjust as needed
    linewidth = 0.9
  )+

  # =================================================================
  # Custom axes: draw axis lines first
  # =================================================================
  # Y-axis: vertical line + left-facing tick marks
  geom_segment(
    inherit.aes = FALSE,
    aes(x = y_axis_x, xend = y_axis_x,
        y = y_axis_min, yend = y_axis_max),
    linewidth = 0.7
  ) +
  geom_segment(
    inherit.aes = FALSE,
    data = data.frame(y = y_breaks_used),
    aes(x = y_tick_x1, xend = y_tick_x2,
        y = y, yend = y),
    linewidth = 0.7
  ) +

  # X-axis: horizontal line + downward tick marks for each category
  geom_segment(
    inherit.aes = FALSE,
    aes(x = x_simple, xend = x_complex,
        y = x_axis_y, yend = x_axis_y),
    linewidth = 0.7
  ) +
  geom_segment(
    inherit.aes = FALSE,
    data = data.frame(x = c(x_simple, x_complex)),
    aes(x = x, xend = x,
        y = x_tick_y1, yend = x_tick_y2),
    linewidth = 0.7
  ) +

  # =================================================================
  # Manually add tick labels and axis titles near the custom axes
  # =================================================================

  # Y tick labels (slightly to the right of the axis)
  geom_text(
    inherit.aes = FALSE,
    data = data.frame(
      y     = y_breaks_used,
      label = y_breaks_used
    ),
    aes(x = y_axis_x - 0.15, y = y, label = label),
    hjust = 0, vjust = 0.5, size = 3.5
  ) +

  # X tick labels (slightly below the downward ticks)
  geom_text(
    inherit.aes = FALSE,
    data = data.frame(
      x     = c(x_simple, x_complex),
      label = c("Simple", "Complex")
    ),
    aes(x = x, y = x_tick_y2 - 0.03 * dy, label = label),
    vjust = 1, size = 3.5
  ) +

  # Y-axis title (centered; rotated 90°)
  geom_text(
    inherit.aes = FALSE,
    aes(x = y_axis_x - 0.3,
        y = (y_axis_min + y_axis_max) / 2,
        label = "HeadMovement (m)"),
    angle = 90, hjust = 0.5, vjust = 0.5, size = 4
  ) +

  # X-axis title (centered; below tick labels)
  geom_text(
    inherit.aes = FALSE,
    aes(x = (x_simple + x_complex) / 2,
        y = x_tick_y2 - 0.10 * dy,
        label = "Complexity"),
    vjust = 1, size = 4
  ) +

  # ===== Colors & legend =====
  scale_colour_manual(
    values = c("3D-only" = col_blue,
               "2D+3D"   = col_red),
    name   = "Condition"
  ) +
  scale_fill_manual(
    values = c("3D-only" = alpha(col_blue, 0.35),
               "2D+3D"   = alpha(col_red,  0.35)),
    name   = "Condition"
  ) +

  # ===== Coordinate scaling (continuous x + custom ticks) =====
  scale_y_continuous(
    limits = c(y_plot_min, y_plot_max),
    breaks = y_breaks_used
  ) +
  scale_x_continuous(
    limits = c(x_plot_min, x_plot_max),
    breaks = c(x_simple, x_complex),
    labels = NULL   # suppress default labels; custom labels are drawn
  ) +

  # ===== Theme =====
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),   # remove major y-grid lines
    panel.grid.minor   = element_blank(),
    axis.text.x        = element_blank(),   # remove default x-axis tick labels
    axis.text.y        = element_blank(),
    axis.title.x       = element_blank(),   # remove default x-axis title
    axis.title.y       = element_blank(),
    axis.ticks         = element_blank(),
    axis.line          = element_blank(),
    legend.position    = "top",
    legend.title       = element_text(size = 11),
    legend.margin      = margin(b = -5),
    plot.margin        = margin(10, 25, 35, 45)
  ) +
  coord_cartesian(clip = "off")

p_ct_raincloud
```


# 4. Light robustness checks (optional)

- **Option A**: if residual heteroscedasticity is still a concern, set `use_varIdent_as_main = FALSE` to compare against `log-LMM (RI)`.  
- **Option B**: **cluster bootstrap CIs (cluster = participant)** (runs when `params$do_cluster_boot = TRUE`; note the computational cost).

```{r eval=params$do_cluster_boot}
cluster_ids <- unique(df$participant)
B <- params$boot_B
boot_est <- replicate(B, {
  ids <- sample(cluster_ids, replace=TRUE)
  d2 <- dplyr::bind_rows(lapply(ids, function(pid) dplyr::filter(df, participant == pid)))
  fit_b <- try(lme4::lmer(log_head_move ~ group*complexity + (1|participant), data=d2, REML=TRUE), silent=TRUE)
  if(inherits(fit_b,"try-error")) return(rep(NA_real_, 4))
  em <- emmeans::emmeans(fit_b, ~ group*complexity) %>% summary()
  as.numeric(em$emmean) # log-EMM for the four cells
})

boot_df <- as.data.frame(t(boot_est))
colnames(boot_df) <- paste0("cell",1:4)
boot_ci <- apply(boot_df, 2, function(x) quantile(exp(x), probs=c(.025,.5,.975), na.rm=TRUE))
as.data.frame(t(boot_ci))
```

# 5. One-off sensitivity analysis: Gamma-GLMM (log link)

```{r}
sens_fit <- glmmTMB::glmmTMB(head_move ~ group*complexity + (1|participant),
                              data=df, family=Gamma(link="log"))
emm_sens <- emmeans::emmeans(sens_fit, ~ group*complexity, type="response")
summary(emm_sens)
```



## 7. Head Movement — Δ EMMs Forest (integrated table + plot, aligned version)

> This block keeps the plotting implementation unchanged (mirrors the Completion Time forest-plot section); only Head Movement results are used as inputs.

```{r hm_forest_aligned, fig.width=10, fig.height=5, out.width='100%'}
suppressPackageStartupMessages({
  library(emmeans); library(dplyr); library(tidyr); library(forcats)
  library(ggplot2); library(scales); library(glue)
})

# === 1) Extract cell-level EMMs (link scale: log seconds) ===
emm_cells <- emmeans(main_fit, ~ group*complexity)

# 2) Back-transform to response scale (seconds): geometric mean = exp(EMM)
emm_link  <- summary(emm_cells, infer = TRUE)

# 3) Build the cell table for A|B contrasts — time_sec must be on the response scale (seconds)
cells_df <- as.data.frame(emm_link) |>
  dplyr::transmute(
    group, complexity,
    time_sec = exp(emmean)
  )

# === 2) Four contrasts (response scale; Holm + CI) ===
grp_res <- contrast(emm_cells, method = "revpairwise", by = "complexity", adjust = "holm") %>%
  summary(infer = TRUE) %>% as.data.frame() %>%
  dplyr::mutate(
    Contrast = dplyr::case_when(
      complexity == "simple"  ~ "Simple Scenario",
      complexity == "complex" ~ "Complex Scenario",
      TRUE ~ paste0("Group@", complexity)
    ),
    Block = "Condition Effect (2D+3D vs. 3D-only) "
  )

cpx_res <- contrast(emm_cells, method = "revpairwise", by = "group", adjust = "holm") %>%
  summary(infer = TRUE) %>% as.data.frame() %>%
  dplyr::mutate(
    Contrast = dplyr::case_when(
      group == "3D-only" ~ "3D-only Condition",
      group == "2D+3D"   ~ "2D+3D Condition",
      TRUE ~ paste0("Complexity@", group)
    ),
    Block = "Complexity Effect (Complex vs. Simple)"
  )

ct_contr <- dplyr::bind_rows(grp_res, cpx_res) %>%
  dplyr::transmute(
    Contrast, Block,
    ln_ratio = estimate, ln_low = lower.CL, ln_high = upper.CL,
    gmr = exp(estimate), gmr_low = exp(lower.CL), gmr_high = exp(upper.CL),
    p_holm = p.value, group, complexity
  )

# === 3) Fixed ordering + row spacing (tunable) ===
row_order <- c("Simple Scenario", "Complex Scenario", "3D-only Condition", "2D+3D Condition")

gap_within_block1 <- 0.20
gap_between_blocks <- 0.60
gap_within_block2 <- 0.20

y1 <- 0.3
y2 <- y1 - gap_within_block1
y3 <- y2 - gap_between_blocks
y4 <- y3 - gap_within_block2

row_y <- c("Simple Scenario"=y1,"Complex Scenario"=y2,"3D-only Condition"=y3,"2D+3D Condition"=y4)
row_map <- tibble::tibble(Contrast = factor(names(row_y), levels = row_order),
                          y = as.numeric(row_y[as.character(Contrast)]))

ct_contr <- ct_contr %>%
  dplyr::mutate(Contrast = factor(Contrast, levels = row_order)) %>%
  dplyr::left_join(row_map, by = "Contrast") %>%
  dplyr::arrange(dplyr::desc(y)) %>%
  dplyr::rowwise() %>%
  dplyr::mutate(
    baseline = dplyr::case_when(
      Contrast == "Simple Scenario"        ~ cells_df$time_sec[cells_df$group=="3D-only" & cells_df$complexity=="simple"],
      Contrast == "Complex Scenario"       ~ cells_df$time_sec[cells_df$group=="3D-only" & cells_df$complexity=="complex"],
      Contrast == "3D-only Condition"  ~ cells_df$time_sec[cells_df$group=="3D-only" & cells_df$complexity=="simple"],
      Contrast == "2D+3D Condition"    ~ cells_df$time_sec[cells_df$group=="2D+3D"   & cells_df$complexity=="simple"],
      TRUE ~ NA_real_
    ),
    a_time = dplyr::case_when(
      Contrast == "Simple Scenario"        ~ cells_df$time_sec[cells_df$group=="2D+3D" & cells_df$complexity=="simple"],
      Contrast == "Complex Scenario"       ~ cells_df$time_sec[cells_df$group=="2D+3D" & cells_df$complexity=="complex"],
      Contrast == "3D-only Condition"  ~ cells_df$time_sec[cells_df$group=="3D-only" & cells_df$complexity=="complex"],
      Contrast == "2D+3D Condition"    ~ cells_df$time_sec[cells_df$group=="2D+3D"   & cells_df$complexity=="complex"],
      TRUE ~ NA_real_
    ),
    pct_change = 100 * (gmr - 1)
  ) %>% dplyr::ungroup()

y_top    <- max(row_map$y)
y_bottom <- min(row_map$y)

# === 4) Left-table column positions ===
col_x <- c(
  "Contrast" = 0.15,
  "EMMs A|B (m)" = 1.45,
  "GMR [95% CI]" = 2.25,
  "p (Holm)" = 3.35
)
xlim_tbl <- c(0.05, 3.6)

# Horizontal position and width of the forest panel within the shared canvas
forest_gap_x <- 0.40
forest_width <- 2.20
x_forest_min <- xlim_tbl[2] + forest_gap_x
x_forest_max <- x_forest_min + forest_width

# === 5) Forest numeric range + 'nice' ticks (prefer 0.5 / 0.25 / 1) ===
xr_raw <- range(c(ct_contr$gmr_low, ct_contr$gmr_high), na.rm = TRUE)
pad    <- max(diff(xr_raw) * 0.06, 0.03)
xlim_for <- c(xr_raw[1] - pad, xr_raw[2] + pad)

choose_step <- function(rng, target_ticks = 4) {
  width <- diff(rng)
  cand  <- c(0.1, 0.2, 0.25, 0.5, 1, 2)
  cand[which.min(abs(width/cand - target_ticks))]
}
step <- choose_step(xlim_for, target_ticks = 4)

low  <- floor(xlim_for[1]/step) * step
high <- ceiling(xlim_for[2]/step) * step
x_breaks_for <- seq(low, high, by = step)
# Ensure 0 is within the range so it always appears
if (xlim_for[1] <= 1 && xlim_for[2] >= 1) x_breaks_for <- sort(unique(c(x_breaks_for, 1)))
x_breaks_for <- x_breaks_for[x_breaks_for >= xlim_for[1] & x_breaks_for <= xlim_for[2]]

# Map numeric values to forest x-coordinates on the shared canvas
xf <- function(x) scales::rescale(x, to = c(x_forest_min, x_forest_max), from = xlim_for)

ct_contr <- ct_contr %>%
  dplyr::mutate(x_low = xf(gmr_low), x_high = xf(gmr_high), x_mid = xf(gmr))

x_ticks <- data.frame(
  brk = x_breaks_for,
  x   = xf(x_breaks_for)
)
tick_lab <- scales::label_number(accuracy = step)(x_ticks$brk)

# === 6) Header / thick rule / section labels (all tunable) ===
header_text_y   <- y_top + 0.65     # header text top position
header_line_gap <- 0.08             # gap from header to thick rule (larger -> rule moves down)
header_line_y   <- header_text_y - header_line_gap

header_df <- data.frame(
  x     = unname(col_x),
  y     = header_text_y,
  label = names(col_x)
)

sec_offset_top    <- 0.20   # Condition Effect (2D+3D vs. 3D-only)  vs Simple Scenario
sec_offset_bottom <- 0.20   # Complexity Effect (Complex vs. Simple) vs 3D-only Condition
sec_labels <- dplyr::tibble(
  label = c("Condition Effect (2D+3D vs. 3D-only) ", "Complexity Effect (Complex vs. Simple)"),
  y     = c(row_y["Simple Scenario"] + sec_offset_top,
            row_y["3D-only Condition"] + sec_offset_bottom),
  x     = unname(col_x["Contrast"])
)
# === 6.9) Shift all rows below the thick rule upwards slightly (without moving the rule) ===
body_dy <- 0.15   # tunable: positive = up, negative = down (e.g., 0.10–0.25)
ct_contr  <- ct_contr  %>% dplyr::mutate(y = y + body_dy)   # rows (left/right synchronized)
sec_labels <- sec_labels %>% dplyr::mutate(y = y + body_dy) # shift section labels as well

# === 7) Draw a custom bottom axis (line + ticks + labels) and clip the 0 dashed line to the axis ===
axis_gap       <- 0.15           # distance from bottom row to axis
axis_y         <- y_bottom - axis_gap
tick_up        <- 0.08
tick_down      <- 0.06
tick_label_gap <- 0.28
tick_label_y   <- axis_y - tick_label_gap

# The y-range must include the axis line, ticks, and labels
ylims_lower <- axis_y - (tick_label_gap + tick_down + 0.18)
ylims_upper <- y_top + 1.90
ylims       <- c(ylims_lower, ylims_upper)
y_breaks    <- row_map$y

# Vertical 0 dashed line: from bottom axis to the thick rule
draw_one <- (xlim_for[1] <= 1 && xlim_for[2] >= 1)
vline_df  <- if (draw_one) data.frame(x = xf(1)) else data.frame(x = numeric(0))

# === 8) Draw on a single canvas ===
tbl_show <- ct_contr %>%
  dplyr::transmute(
    Contrast, y,
    `EMMs A|B (m)` = sprintf("%s | %s", format(round(a_time, 2), nsmall = 2), format(round(baseline, 2), nsmall = 2)),
    `GMR [95% CI]` = sprintf("%.3f [%.3f, %.3f]", gmr, gmr_low, gmr_high),
    `p (Holm)` = ifelse(p_holm < .001, "< .001", format(round(p_holm, 3), nsmall = 3))
  )
tbl_long <- tbl_show %>%
  tidyr::pivot_longer(cols = c("Contrast","EMMs A|B (m)","GMR [95% CI]","p (Holm)"),
                      names_to = "col", values_to = "val") %>%
  dplyr::mutate(x = unname(col_x[col]), is_bold = (col == "Contrast"))

ggplot() +
  # Left: section labels + table body + header
  geom_text(data = sec_labels, aes(x = x, y = y, label = label),
            hjust = 0, vjust = 0, fontface = "bold", size = 4) +
  geom_text(data = tbl_long,
            aes(x = x, y = y, label = val, fontface = ifelse(is_bold,"bold","plain"), colour = "black"),
            hjust = 0, vjust = 0.5, size = 3.6, lineheight = 1.0) +
  geom_text(data = header_df, aes(x = x, y = y, label = label),
            hjust = 0, vjust = 0, size = 3.7, fontface = "bold") +
  scale_colour_identity() +

  # Thick rule spanning the full width
  geom_segment(aes(x = -Inf, xend = Inf, y = header_line_y, yend = header_line_y),
               linewidth = 1.0) +

  # Right: forest CIs + point estimates
  geom_segment(data = ct_contr, aes(x = x_low, xend = x_high, y = y, yend = y), linewidth = 0.9) +
  geom_point(data = ct_contr, aes(x = x_mid, y = y), size = 2.8) +

  # 0 dashed line (from axis to thick rule)
  geom_segment(data = vline_df,
               aes(x = x, xend = x), y = axis_y, yend = header_line_y,
               linetype = "dashed", linewidth = 0.6) +

  # Forest top title (same y as the left-table header)
  annotate("text", x = x_forest_min, y = header_text_y, label = "GMR [95% CI]",
           hjust = -1.1, vjust = 0, size = 3.7, fontface = "bold") +

  # — Custom bottom axis (ticks first, then axis line; redraw axis line last to ensure visibility) —
  geom_segment(data = x_ticks,
               aes(x = x, xend = x, y = axis_y - tick_down, yend = axis_y + tick_up),
               linewidth = 0.6, lineend = "butt") +
  geom_text(data = x_ticks, aes(x = x, y = tick_label_y, label = tick_lab),
            vjust = 1, size = 3.2) +
  annotate("segment", x = x_forest_min, xend = x_forest_max,
           y = axis_y, yend = axis_y, linewidth = 1.0, lineend = "square") +

  # Coordinates and theme
  coord_cartesian(xlim = c(xlim_tbl[1], x_forest_max + 0.05), ylim = ylims, clip = "on") +
  theme_minimal(base_size = 12) +
  theme(
    axis.title  = element_blank(),
    axis.text   = element_blank(),
    panel.grid  = element_blank(),
    plot.margin = margin(t = 6, r = 6, b = 6, l = 4)
  )
```

