---
title: "Click strategy deviation (strategy_div): Model selection and evaluation (single-stage ZIB vs two-part hurdle)"
author: "CongLiu"
params:
  data_path: "Experimental_Raw_Data.csv"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    word_document: default
    pdf_document: default
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(message = FALSE, warning = FALSE,
                      fig.width = 7, fig.height = 4.8,
                      fig.align = "center", dpi = 300,
                      out.width = "90%")
set.seed(2025)
```

# 0. Packages
```{r}
need <- c("tidyverse","janitor","glmmTMB","DHARMa","emmeans",
          "performance","broom.mixed","kableExtra","patchwork")
to_install <- setdiff(need, rownames(installed.packages()))
if(length(to_install))
  install.packages(to_install, repos = "https://cloud.r-project.org")

suppressPackageStartupMessages({
  library(tidyverse);  library(janitor);   library(glmmTMB)
  library(DHARMa);     library(emmeans);   library(performance)
  library(broom.mixed);library(kableExtra);library(patchwork)
})

# Helper functions
coalesce_col <- function(df, patterns, new){
  hit <- names(df)[stringr::str_detect(
    names(df),
    stringr::regex(paste0("(", paste(patterns, collapse="|"), ")"),
                   ignore_case = TRUE))]
  if(length(hit)==0) return(df)
  dplyr::rename(df, !!new := dplyr::all_of(hit[1]))
}

# Fail-safe model fitting: capture errors and return NULL
fit_safely <- function(expr)
  tryCatch(expr, error = function(e){
    message("⚠️ ", conditionMessage(e)); return(NULL)
  })

# Always return a comparable AIC value (Inf if unavailable)
safe_AIC <- function(mod){
  if(is.null(mod)) return(Inf)                         # fit failed
  out <- tryCatch(AIC(mod), error = function(e) NA_real_)
  if(is.na(out)) Inf else out                          # log-likelihood unavailable
}

# ---- Data path resolution (recommended for reproducibility) ----
resolve_data_path <- function(path,
                              candidate_dirs = c("..", "../data", "../datasets", "../dataset",
                                                 "../raw", "../raw_data", "../input", "../inputs"),
                              must_exist = TRUE) {
  if (!is.null(path) && nzchar(path) && file.exists(path)) {
    return(normalizePath(path, winslash = "/", mustWork = FALSE))
  }
  fname <- basename(path)
  candidates <- unique(c(fname, file.path(candidate_dirs, fname)))
  hit <- candidates[file.exists(candidates)][1]
  if (is.na(hit) || is.null(hit)) {
    msg <- paste0(
      "Data file not found. Tried the following locations:\n",
      paste(candidates, collapse = "\n")
    )
    if (must_exist) stop(msg, call. = FALSE) else return(NA_character_)
  }
  normalizePath(hit, winslash = "/", mustWork = FALSE)
}

data_file <- resolve_data_path(params$data_path)

```

# 1. Data import and preprocessing
```{r}
raw <- readr::read_csv(data_file, show_col_types = FALSE) %>% 
         janitor::clean_names()

df <- raw %>% 
  coalesce_col(c("^participant$","^subj","^pid","^ppid"), "participant") %>% 
  coalesce_col(c("^group$","condition","interact","arm"), "group") %>% 
  coalesce_col(c("^complexity$","path_complex","task_complexity",
                 "simple|complex"), "complexity") %>% 
  coalesce_col(c("^strategy_?div$","click.*strategy","kendall",
                 "tau","order.*dev"), "strategy_div") %>% 
  mutate(
    participant = as.factor(participant),
    group = forcats::fct_relabel(as.factor(group), \(x)
              ifelse(str_detect(x, regex("2D", ignore_case = TRUE)),
                     "2D+3D","3D-only")) |> forcats::fct_relevel("3D-only","2D+3D"),
    complexity = forcats::fct_relabel(as.factor(complexity), \(x)
              ifelse(str_detect(x, regex("complex", ignore_case = TRUE)),
                     "complex","simple")) |> forcats::fct_relevel("simple","complex"),
    strategy_div = as.numeric(strategy_div)
  )

df$y0  <- as.integer(df$strategy_div == 0)
pos    <- filter(df, strategy_div > 0)  # positive subset
```

# 2. Candidate models
- **Single-stage (ZIB)**: `beta_family(link="logit")` + `ziformula`  
- **Two-part (hurdle)**: (i) zero vs non-zero (binomial), then (ii) positive values (Gamma/Beta).

# 3. Model fitting (with fail-safe wrappers)
```{r}
ctrl <- glmmTMBControl(optimizer = optim, optArgs = list(method = "BFGS"))

# --- Single-stage ZIB ---
ZIB_int   <- fit_safely(
  glmmTMB(strategy_div ~ group*complexity + (1|participant),
          ziformula   = ~ group*complexity + (1|participant),
          dispformula = ~ 1,
          data = df, family = beta_family(link="logit"),
          control = ctrl))

ZIB_slope <- fit_safely(
  glmmTMB(strategy_div ~ group*complexity + (1+complexity|participant),
          ziformula   = ~ group*complexity + (1|participant),
          dispformula = ~ 1,
          data = df, family = beta_family(link="logit"),
          control = ctrl))

# --- Two-part: zero vs non-zero ---
L1 <- fit_safely(
        glmmTMB(y0 ~ group*complexity + (1|participant),
                data=df, family = binomial(), control=ctrl))
L2 <- fit_safely(
        glmmTMB(y0 ~ group*complexity + (1+complexity|participant),
                data=df, family = binomial(), control=ctrl))

# --- Two-part: positive component ---
M1b <- fit_safely(
        glmmTMB(strategy_div ~ group*complexity + (1|participant),
                data=pos, family = Gamma(link="log"), control=ctrl))
M1c <- fit_safely(
        glmmTMB(strategy_div ~ group*complexity + (1+complexity|participant),
                data=pos, family = Gamma(link="log"), control=ctrl))

M2a <- fit_safely(
        glmmTMB(strategy_div ~ group*complexity + (1|participant),
                data=pos, family = beta_family(link="logit"), control=ctrl))
M2b <- fit_safely(
        glmmTMB(strategy_div ~ group*complexity + (1+complexity|participant),
                data=pos, family = beta_family(link="logit"), control=ctrl))
```

# 4. Information criteria and diagnostics
```{r}
# --- AIC / BIC summary ---
mods_single <- list(ZIB_int=ZIB_int, ZIB_slope=ZIB_slope)
mods_zero   <- list(L1=L1, L2=L2)
mods_pos    <- list(M1b=M1b, M1c=M1c, M2a=M2a, M2b=M2b)

grab_ic <- function(lst){
  map_dfr(names(lst), \(nm){
    m <- lst[[nm]]
    tibble(model = nm,
           AIC  = if(is.null(m)) NA_real_ else AIC(m),
           BIC  = if(is.null(m)) NA_real_ else BIC(m))
  })
}

aic_single <- grab_ic(mods_single)
aic_zero   <- grab_ic(mods_zero)
aic_pos    <- grab_ic(mods_pos)

# Optional: convert Inf to NA for presentation
show_tbl <- function(x, cap){
  kable(x %>% mutate(across(c(AIC,BIC), ~ ifelse(is.infinite(.), NA, .))),
        digits = 2, caption = cap) %>%
    kable_styling(full_width = FALSE)
}
show_tbl(aic_single, "Single-stage ZIB candidates (AIC/BIC)")
show_tbl(aic_zero,   "Two-part: zero vs non-zero (binomial) candidates (AIC/BIC)")
show_tbl(aic_pos,    "Two-part: positive component (Gamma/Beta) candidates (AIC/BIC)")
```

```{r}
# --- DHARMa residual diagnostics (example; optional) ---
safe_plot_res <- function(mod){
  if(!is.null(mod))
    try(plot(DHARMa::simulateResiduals(mod, n = 300, plot = FALSE)),
        silent = TRUE)
}
safe_plot_res(ZIB_int)
safe_plot_res(ZIB_slope)
walk(list(M1b, M1c, M2a, M2b), safe_plot_res)
```

# 5. Automated recommendation
```{r}
# --- Single-stage ZIB ---
best_ZIB  <- if(safe_AIC(ZIB_slope) < safe_AIC(ZIB_int))
               "ZIB_slope" else "ZIB_int"

# --- Zero vs non-zero ---
best_zero <- if(safe_AIC(L2) < safe_AIC(L1)) "L2" else "L1"

# --- Positive component ---
cand_pos  <- list(M2b = M2b, M2a = M2a, M1c = M1c, M1b = M1b)
scores    <- purrr::map_dbl(cand_pos, safe_AIC)

best_pos <- if (all(is.infinite(scores))) {
  NA_character_
} else {
  names(scores)[which.min(scores)]
}

tibble(best_ZIB = best_ZIB,
       best_zero = best_zero,
       best_pos = best_pos) %>%
  kable(caption = "Automated recommendation (AIC-based heuristic)") %>%
  kable_styling(full_width = FALSE)

```{r SD_diag, fig.width=10, fig.height=12, out.width="100%", dpi=300}
performance::check_model(ZIB_int)
```
