
---
title: "Inference and Post-hoc Analyses (Outcome: strategy_div; ZIB one-step + two-step synthesis)"
author: "CongLiu"
params:
  data_path: "Experimental_Raw_Data.csv"
  B: 3000   # Number of parameter draws for Monte Carlo uncertainty (two-step synthesis)
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  comment = NA,
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  results = "asis",
  fig.align = "center",
  fig.width = 7,
  fig.height = 4.8,
  dpi = 300,
  fig.retina = 2
)
set.seed(2025)

options(width = 120)
options(knitr.kable.NA = "")

# Helper: explicitly print kable objects so knitr captures the rendered tables
kbl_show <- function(x){
  print(x)
  invisible(x)
}

# Helper: resolve the data file path when the raw data are stored *outside* the
# current analysis folder (e.g., the data file sits next to this folder).
#
# Edit `candidate_dirs` below if your project uses a different layout.
resolve_data_path <- function(
  filename,
  candidate_dirs = c(
    "..",
    "../data", "../datasets", "../raw", "../input",
    "../Data", "../DATA",
    "../..", "../../data", "../../datasets",
    ".",  # fallback (legacy): same folder as this Rmd
    "./data"
  )
){
  paths <- unique(file.path(candidate_dirs, filename))
  hit <- paths[file.exists(paths)]
  if(length(hit) > 0) return(hit[1])

  stop(
    paste0(
      "Data file not found: ", filename, "
",
      "Searched the following locations (in order):
",
      paste0(" - ", normalizePath(paths, winslash = "/", mustWork = FALSE), collapse = "
")
    )
  )
}
```

# Dependencies and data
```{r packages}
need <- c("tidyverse","janitor","glmmTMB","emmeans","MASS","parameters","kableExtra","patchwork","gghalves","pBrackets")
to_install <- setdiff(need, rownames(installed.packages()))
if(length(to_install))
  install.packages(to_install, repos = "https://cloud.r-project.org")

suppressPackageStartupMessages({
  library(tidyverse); library(janitor); library(glmmTMB)
  library(emmeans);    library(MASS);   library(kableExtra)
  library(patchwork);  library(gghalves); 
  library(parameters); library(pBrackets)
})

## ---------- Utility functions ----------
coalesce_col <- function(df, patterns, new){
  hit <- names(df)[stringr::str_detect(
    names(df),
    stringr::regex(paste0("(", paste(patterns, collapse="|"), ")"),
                   ignore_case = TRUE))]
  if(length(hit)==0) return(df)
  dplyr::rename(df, !!new := dplyr::all_of(hit[1]))
}

safe_AIC <- function(mod){
  if(is.null(mod) || inherits(mod, "try-error")) return(Inf)
  out <- tryCatch(AIC(mod), error = function(e) NA_real_)
  if(is.na(out)) Inf else out
}
```


```{r}
# ---------- Load and clean data ----------
data_file <- resolve_data_path(params$data_path)
raw <- readr::read_csv(data_file, show_col_types = FALSE) %>%
         janitor::clean_names()

df <- raw %>%
  coalesce_col(c("^participant$","participantid","^subject$","subject",
                 "^subj$","^pid$","^ppid$"), "participant") %>%
  coalesce_col(c("^group$","condition","interact","arm"), "group") %>%
  coalesce_col(c("^complexity$","path_complex","task_complexity","simple|complex"), "complexity") %>%
  coalesce_col(c("^strategy_?div$","click.*strategy","kendall","tau","order.*dev"), "strategy_div")

## If 'participant' is still missing, create a placeholder column
if(!"participant" %in% names(df)){
  warning("Participant ID column not found; created a dummy 'participant' column.")
  df$participant <- seq_len(nrow(df))
}

## Type conversion and basic validation
df <- df %>%
  mutate(
    participant = factor(participant),
    group = forcats::fct_relabel(factor(group), \(x)
              ifelse(stringr::str_detect(x, regex("2D", ignore_case = TRUE)),
                     "2D+3D","3D-only")) |> forcats::fct_relevel("3D-only","2D+3D"),
    complexity = forcats::fct_relabel(factor(complexity), \(x)
              ifelse(stringr::str_detect(x, regex("complex", ignore_case = TRUE)),
                     "complex","simple")) |> forcats::fct_relevel("simple","complex"),
    strategy_div = as.numeric(strategy_div)
  )

df$y0 <- as.integer(df$strategy_div == 0)
pos   <- dplyr::filter(df, strategy_div > 0)
```

# Model fitting: one-step ZIB and two-step approach
```{r}
ctrl <- glmmTMBControl(optimizer = optim,
                       optArgs  = list(method = "BFGS"))

# -- One-step ZIB: use the best model selected in 02 (best_ZIB = "ZIB_int")
ZIB_best <- try(
  glmmTMB(strategy_div ~ group*complexity + (1|participant),
          ziformula   = ~ group*complexity + (1|participant),
          dispformula = ~ 1,
          data = df, family = beta_family(link="logit"), control = ctrl),
  silent = TRUE)

if(inherits(ZIB_best, "try-error")){
  warning("The selected ZIB model did not converge; one-step ZIB outputs will be skipped.")
  ZIB_best <- NULL
}

# -- Two-step: zero/non-zero component; best_zero = "L1" (from 02)
L_best <- try(
  glmmTMB(y0 ~ group*complexity + (1|participant),
          data = df, family = binomial(), control = ctrl),
  silent = TRUE)
if(inherits(L_best, "try-error"))
  stop("The zero/non-zero (hurdle) component did not converge.")

# -- Two-step: positive component; best_pos = "M2a" (Beta, random intercept; from 02)
P_best <- try(
  glmmTMB(strategy_div ~ group*complexity + (1|participant),
          data = pos, family = beta_family(link="logit"), control = ctrl),
  silent = TRUE)

if(inherits(P_best, "try-error")){
  message("M2a (Beta) 未收敛，退回 Gamma(log) 同一随机结构。")
  P_best <- glmmTMB(strategy_div ~ group*complexity + (1|participant),
                    data = pos, family = Gamma(link="log"), control = ctrl)
  pos_family <- "Gamma"
} else {
  pos_family <- "Beta"
}
```

## Tab 2 — Fixed-effects summary（Main models）

```{r tab2_fixef_zib, results='asis', message=FALSE, warning=FALSE}
fe_table <- function(fit, component=NULL, reference="3D-only × simple", digits=3){
  if (!requireNamespace("parameters", quietly = TRUE)) install.packages("parameters", repos="https://cloud.r-project.org")
  suppressPackageStartupMessages({ library(parameters); library(dplyr) })
  if(inherits(fit, "glmmTMB")){
    comp <- if(is.null(component)) "conditional" else component
    tab <- parameters::model_parameters(fit, component = comp, ci_method = "wald", verbose = FALSE)
  } else {
    tab <- parameters::model_parameters(fit, ci_method = "wald", verbose = FALSE)
  }
  stat_col <- if("z" %in% names(tab)) "z" else if("t" %in% names(tab)) "t" else if("Statistic" %in% names(tab)) "Statistic" else names(tab)[1]
  out <- tab %>% dplyr::transmute(
    Term = Parameter, `β` = Coefficient, SE = SE,
    Stat = .data[[stat_col]], p = p,
    `95% CI` = sprintf("[%.3f, %.3f]", CI_low, CI_high)
  )
  kableExtra::kable(out, digits = digits,
    caption = paste0("Fixed-effects — ", deparse(substitute(fit)),
                     if(!is.null(component)) paste0(" [", component, "]") else "", 
                     " — reference: ", reference)) %>% 
    kableExtra::kable_styling(full_width = FALSE)
}


# ========== Fixed Effects Table of Overall (E[y]) (Main Effects × Interaction) ==========
fe_table_overall <- function(fit, reference="3D-only × simple", digits=3, include_intercept = TRUE){
  if (is.null(fit)) return(invisible(NULL))
  suppressPackageStartupMessages({ library(dplyr); library(knitr); library(kableExtra) })

  cells <- expand.grid(group = c("3D-only","2D+3D"),
                       complexity = c("simple","complex")) %>%
    dplyr::mutate(group      = factor(group,      levels = c("3D-only","2D+3D")),
                  complexity = factor(complexity, levels = c("simple","complex")),
                  cell       = paste(group, complexity, sep = "·"))

  pr0 <- predict(fit, newdata = cells, type = "zprob",       se.fit = TRUE, re.form = NA)
  mu  <- predict(fit, newdata = cells, type = "conditional", se.fit = TRUE, re.form = NA)

  Ey   <- (1 - pr0$fit) * mu$fit
  seEy <- sqrt(((1 - pr0$fit)^2) * (mu$se.fit^2) + (mu$fit^2) * (pr0$se.fit^2))

  A_mean <- Ey[cells$cell == "3D-only·simple"];  A_se <- seEy[cells$cell == "3D-only·simple"]
  B_mean <- Ey[cells$cell == "3D-only·complex"]; B_se <- seEy[cells$cell == "3D-only·complex"]
  C_mean <- Ey[cells$cell == "2D+3D·simple"];    C_se <- seEy[cells$cell == "2D+3D·simple"]
  D_mean <- Ey[cells$cell == "2D+3D·complex"];   D_se <- seEy[cells$cell == "2D+3D·complex"]

  beta_group    <- ((C_mean + D_mean)/2) - ((A_mean + B_mean)/2)
  se_group      <- sqrt((C_se^2 + D_se^2 + A_se^2 + B_se^2) / 4)

  beta_complex  <- ((B_mean + D_mean)/2) - ((A_mean + C_mean)/2)
  se_complex    <- sqrt((B_se^2 + D_se^2 + A_se^2 + C_se^2) / 4)

  beta_inter    <- (D_mean - B_mean) - (C_mean - A_mean)
  se_inter      <- sqrt(D_se^2 + B_se^2 + C_se^2 + A_se^2)

  out <- dplyr::tibble(
    Term   = c("Group (2D+3D vs 3D-only)",
               "Complexity (complex vs simple)",
               "Interaction"),
    `β`    = c(beta_group, beta_complex, beta_inter),
    SE     = c(se_group,   se_complex,   se_inter),
    Stat   = c(beta_group/se_group, beta_complex/se_complex, beta_inter/se_inter),
    p      = 2*pnorm(abs(c(beta_group/se_group, beta_complex/se_complex, beta_inter/se_inter)), lower.tail = FALSE),
    `95% CI` = sprintf("[%.3f, %.3f]",
                       c(beta_group, beta_complex, beta_inter) - 1.96*c(se_group, se_complex, se_inter),
                       c(beta_group, beta_complex, beta_inter) + 1.96*c(se_group, se_complex, se_inter))
  )

  if (isTRUE(include_intercept)) {
    intercept_row <- dplyr::tibble(
      Term = "(Intercept) = E[y | 3D-only × simple]",
      `β` = A_mean, SE = A_se,
      Stat = A_mean / A_se,
      p = 2*pnorm(abs(A_mean / A_se), lower.tail = FALSE),
      `95% CI` = sprintf("[%.3f, %.3f]", A_mean - 1.96*A_se, A_mean + 1.96*A_se)
    )
    out <- dplyr::bind_rows(intercept_row, out)
  }

  kableExtra::kable(out, digits = digits,
    caption = "Single-step ZIB (Overall: unconditional E[y]): Wald tests for main effects and interactions of Group/Complexity (and optional intercept) (Source: mean difference / difference-in-differences of 2x2 E[y])") %>%
    kableExtra::kable_styling(full_width = FALSE) %>% print()
  invisible(out)
}

if(!is.null(ZIB_best)){
  fe_table_overall(ZIB_best)
  fe_table(ZIB_best, component="zi")
}
if(exists("L_best")) fe_table(L_best, component="conditional")
if(exists("P_best")) fe_table(P_best, component="conditional")
```


## Forest data — standardized effects (Std. β)

```{r forest_std_zib, results='asis', message=FALSE, warning=FALSE}
# Required packages
pkgs <- c("parameters","effectsize","insight","dplyr","knitr")
for(p in pkgs) if(!requireNamespace(p, quietly=TRUE)) install.packages(p, repos="https://cloud.r-project.org")
suppressPackageStartupMessages({ library(parameters); library(effectsize); library(insight); library(dplyr); library(knitr) })

std_table <- function(fit, terms = c("group2D+3D","complexitycomplex","group2D+3D:complexitycomplex"), component=NULL){
  # Try refit-based standardization first
  tab <- NULL
  try({
    if(inherits(fit,"glmmTMB")){
      tab <- effectsize::standardize_parameters(fit, method="refit",
                                                component=if(is.null(component)) "conditional" else component, ci=0.95)
    } else {
      tab <- effectsize::standardize_parameters(fit, method="refit", ci=0.95)
    }
  }, silent = TRUE)

  recode_label <- c("group2D+3D"="Group (2D+3D vs 3D-only)",
                    "complexitycomplex"="Complexity (complex vs simple)",
                    "group2D+3D:complexitycomplex"="Interaction")

  use_refit <- !is.null(tab)
  if(use_refit){
    df <- as.data.frame(tab)
    # Keep only requested terms; if Std_Coefficient is all NA, fall back
    if(all(is.na(df$Std_Coefficient[df$Parameter %in% terms]))){
      use_refit <- FALSE
    } else {
      df <- df %>% 
        rename(
          CI_low = ifelse("CI_low_Std" %in% colnames(.), "CI_low_Std", "CI_low"),
          CI_high = ifelse("CI_high_Std" %in% colnames(.), "CI_high_Std", "CI_high")
        )
      out <- df[df$Parameter %in% terms, , drop=FALSE]
      out <- out %>%
        dplyr::transmute(Effect = dplyr::recode(Parameter, !!!recode_label),
                         `Std. β` = Std_Coefficient,
                         `CI low`  = CI_low,  
                         `CI high` = CI_high) 
      return(out)
    }
  }

  # Fallback path: Wald CI on model scale -> standardize via beta * SD(col in model matrix) / SD(Y)
  raw <- parameters::model_parameters(fit,
          component = if(inherits(fit,"glmmTMB")) "conditional" else NULL,
          ci_method = "wald")
  raw <- as.data.frame(raw)
  raw <- raw[raw$Parameter %in% terms, , drop=FALSE]

  # SD(Y) from data
  mf <- insight::get_data(fit, verbose=FALSE)
  resp <- insight::find_response(fit)
  SDy <- stats::sd(mf[[resp]], na.rm=TRUE)

  # Use *model matrix* so column names match coefficient labels
  mm <- try(insight::get_modelmatrix(fit, component = "conditional"), silent = TRUE)
  if(inherits(mm, "try-error") || is.null(mm)){
    mm <- try(stats::model.matrix(stats::formula(fit), data = mf), silent = TRUE)
  }

  get_sdx <- function(param){
    if(!is.null(mm) && !inherits(mm, "try-error") && param %in% colnames(mm)){
      return(stats::sd(mm[, param], na.rm=TRUE))
    }
    # Last resort: approximate using predictor variance if column can't be found
    varname <- sub(":.*", "", sub("(.*?)(\\w+)$", "\\1\\2", param))
    if(!is.null(mf[[varname]])){
      x <- mf[[varname]]
      ux <- stats::na.omit(unique(x))
      if(length(ux)==2 && all(sort(ux) %in% c(0,1))) {
        p <- mean(x, na.rm=TRUE); return(sqrt(p*(1-p)))
      }
      return(stats::sd(x, na.rm=TRUE))
    }
    return(NA_real_)
  }

  sdx <- vapply(raw$Parameter, get_sdx, numeric(1))
  out <- raw %>%
    dplyr::transmute(Effect = dplyr::recode(Parameter, !!!recode_label),
                     `Std. β` = Coefficient * sdx / SDy,
                     `CI low` = CI_low     * sdx / SDy,
                     `CI high`= CI_high    * sdx / SDy)
  return(out)
}


# ========== The standardized effect of Overall (E[y]) ==========
std_table_overall <- function(fit, dv_name = "strategy_div") {
  stopifnot(!is.null(fit))
  suppressPackageStartupMessages({ library(dplyr); library(tibble) })

  have_outZ <- exists("outZ", envir = .GlobalEnv)
  if (isTRUE(have_outZ)) {
    oz <- get("outZ", envir = .GlobalEnv)
    need_cols <- c("cell","E[y]","lo","hi")
    if (!all(need_cols %in% names(oz))) have_outZ <- FALSE
  }

  if (isTRUE(have_outZ)) {
    cells <- oz %>% dplyr::arrange(match(cell, c("3D-only·simple","3D-only·complex","2D+3D·simple","2D+3D·complex")))
    Ey   <- cells[["E[y]"]]
    seEy <- (cells$hi - cells$lo) / (2*1.96)
    names(Ey)   <- cells$cell
    names(seEy) <- cells$cell
  } else {
    cells <- expand.grid(group = c("3D-only","2D+3D"),
                         complexity = c("simple","complex"),
                         KEEP.OUT.ATTRS = FALSE, stringsAsFactors = FALSE)
    cells$group      <- factor(cells$group,      levels = c("3D-only","2D+3D"))
    cells$complexity <- factor(cells$complexity, levels = c("simple","complex"))
    cells$cell       <- paste(cells$group, cells$complexity, sep = "·")

    pr0 <- predict(fit, newdata = cells, type = "zprob",       se.fit = TRUE, re.form = NA)
    mu  <- predict(fit, newdata = cells, type = "conditional", se.fit = TRUE, re.form = NA)
    Ey   <- (1 - pr0$fit) * mu$fit
    seEy <- sqrt(((1 - pr0$fit)^2) * (mu$se.fit^2) + (mu$fit^2) * (pr0$se.fit^2))
    names(Ey)   <- cells$cell
    names(seEy) <- cells$cell
  }

  if (anyNA(seEy)) {
    seEy[is.na(seEy)] <- 0
  }

  A_mean <- Ey["3D-only·simple"];  A_se <- seEy["3D-only·simple"]
  B_mean <- Ey["3D-only·complex"]; B_se <- seEy["3D-only·complex"]
  C_mean <- Ey["2D+3D·simple"];    C_se <- seEy["2D+3D·simple"]
  D_mean <- Ey["2D+3D·complex"];   D_se <- seEy["2D+3D·complex"]

  beta_group    <- ((C_mean + D_mean)/2) - ((A_mean + B_mean)/2)
  se_group      <- sqrt((C_se^2 + D_se^2 + A_se^2 + B_se^2) / 4)
  beta_complex  <- ((B_mean + D_mean)/2) - ((A_mean + C_mean)/2)
  se_complex    <- sqrt((B_se^2 + D_se^2 + A_se^2 + C_se^2) / 4)
  beta_inter    <- (D_mean - B_mean) - (C_mean - A_mean)
  se_inter      <- sqrt(D_se^2 + B_se^2 + C_se^2 + A_se^2)

  betas <- c(beta_group, beta_complex, beta_inter)
  ses   <- c(se_group,   se_complex,   se_inter)

  if (!exists("df", envir = .GlobalEnv)) stop("Cannot find the cleaned df")
  data_for_sd <- get("df", envir = .GlobalEnv) %>%
    mutate(group = factor(group, levels = c("3D-only","2D+3D")),
           complexity = factor(complexity, levels = c("simple","complex")))

  if (!dv_name %in% names(data_for_sd)) stop(sprintf("Cannot find the DV column in the df '%s'", dv_name))
  y_vec <- suppressWarnings(as.numeric(data_for_sd[[dv_name]]))
  SDy   <- stats::sd(y_vec, na.rm = TRUE)
  if (!is.finite(SDy) || SDy == 0) stop("SD(y) used for standardization is invalid (NA/Inf/0), please check the DV column")

  # Model matrix
  mm <- stats::model.matrix(~ group*complexity, data = data_for_sd)
  want <- c("group2D+3D","complexitycomplex","group2D+3D:complexitycomplex")
  miss <- setdiff(want, colnames(mm))
  if (length(miss)) stop("The matrix is missing a column：", paste(miss, collapse = ", "))

  sdx <- apply(mm[, want, drop = FALSE], 2, stats::sd, na.rm = TRUE)
  if (any(!is.finite(sdx))) stop("SD(x) results in NA/Inf")

  mult <- as.numeric(sdx) / SDy   

  labels <- c("Group (2D+3D vs 3D-only)",
              "Complexity (complex vs simple)",
              "Interaction")

  out <- tibble(
    Effect   = labels,
    `Std. β` = as.numeric(betas) * mult,
    `CI low`  = (as.numeric(betas) - 1.96 * ses) * mult,
    `CI high` = (as.numeric(betas) + 1.96 * ses) * mult
  )
  out$Effect <- as.character(out$Effect)
  out
}

print(std_table_overall(ZIB_best, dv_name = "strategy_div"))

mm <- model.matrix(~ group*complexity, data = df)
print(apply(mm[, c("group2D+3D","complexitycomplex","group2D+3D:complexitycomplex")], 2, sd, na.rm = TRUE))
print(sd(df$strategy_div, na.rm = TRUE))

# --- Forest (merge zi& overall) ---
forest_df_zi <- std_table(ZIB_best, component = "zi") %>%
  dplyr::select(Effect, `Std. β`, `CI low`, `CI high`) %>%
  dplyr::mutate(Component = "zi (Pr[y=0])",
                Effect = as.character(Effect))

forest_df_overall <- std_table_overall(ZIB_best, dv_name = "strategy_div") %>%
  dplyr::mutate(Component = "overall (E[y])",
                Effect = as.character(Effect))

forest_df <- dplyr::bind_rows(forest_df_overall, forest_df_zi) %>%
  dplyr::select(Component, Effect, `Std. β`, `CI low`, `CI high`)

knitr::kable(
  forest_df, digits = 3,
  caption = "Forest data — standardized effects (Std. β with 95% CI; components: zi = Pr[y=0], overall = unconditional E[y]; baseline = 3D-only × simple)"
)
```

# ZIB：Pr(0),E[y\|y>0],E[y]
```{r}
cells <- expand.grid(group = c("3D-only","2D+3D"),
                     complexity = c("simple","complex")) %>%
  mutate(group      = factor(group,      levels = c("3D-only","2D+3D")),
         complexity = factor(complexity, levels = c("simple","complex")),
         cell       = paste(group, complexity, sep = "·"))

if(!is.null(ZIB_best)){
  # ---- 1) Indicator estimation：Pr(y = 0)、E[y|y>0] 、 E[y] ----
  pr0 <- predict(ZIB_best, newdata = cells,
                 type = "zprob",       se.fit = TRUE, re.form = NA)
  mu  <- predict(ZIB_best, newdata = cells,
                 type = "conditional", se.fit = TRUE, re.form = NA)

  Ey   <- (1 - pr0$fit) * mu$fit
  seEy <- sqrt(((1 - pr0$fit)^2) * (mu$se.fit^2) +
               (mu$fit^2)        * (pr0$se.fit^2))

  outZ <- tibble(
    cell        = cells$cell,
    `Pr(y=0)`   = pr0$fit,
    `E[y|y>0]`  = mu$fit,
    `E[y]`      = Ey,
    lo          = Ey - 1.96 * seEy,
    hi          = Ey + 1.96 * seEy
  )

  kbl_show(
    knitr::kable(outZ, digits = 3,
      caption = "Single-Step ZIB: Four-Condition Estimation（Pr(y=0)、E[y|y>0]、E[y]）") %>%
      kableExtra::kable_styling(full_width = FALSE)
  )

  # ---- 2) Overall Effect Test  Pairwise Comparison (ZIB Single-Step Model)----
  emm_ZIB_cond <- try(
    emmeans(ZIB_best, ~ group*complexity,
            component = "cond", type = "response"),
    silent = TRUE)

  if(!inherits(emm_ZIB_cond, "try-error")){
    joint_ZIB_cond <- joint_tests(emm_ZIB_cond)

    kbl_show(
      knitr::kable(as.data.frame(joint_ZIB_cond), digits = 3,
            caption = "Single-step ZIB (cond component): overall Wald test for group/complexity/interaction (E[y|y>0])") %>%
        kableExtra::kable_styling(full_width = FALSE)
    )

    kbl_show(
      knitr::kable(summary(emm_ZIB_cond), digits = 3,
            caption = "Single-step ZIB (cond component): Four-condition marginal means (E[y|y>0], reference cell = 3D-only·simple)") %>%
        kableExtra::kable_styling(full_width = FALSE)
    )

    kbl_show(
      knitr::kable(summary(pairs(emm_ZIB_cond)), digits = 3,
            caption = "Single-step ZIB (cond component): pairwise comparison of four conditions (E[y|y>0])") %>%
        kableExtra::kable_styling(full_width = FALSE)
    )
  } else {
    cat("The current version of emmeans does not support component = 'cond', and only reports point estimates of E[y|y>0]。\n")
  }

  emm_ZIB_zero <- try(
    emmeans(ZIB_best, ~ group*complexity,
            component = "zi", type = "response"),
    silent = TRUE)

  if(!inherits(emm_ZIB_zero, "try-error")){
    joint_ZIB_zero <- joint_tests(emm_ZIB_zero)

    kbl_show(
      knitr::kable(as.data.frame(joint_ZIB_zero), digits = 3,
            caption = "Single-step ZIB (zi component): overall Wald test for group/complexity/interaction (Pr(y=0))") %>%
        kableExtra::kable_styling(full_width = FALSE)
    )

    kbl_show(
      knitr::kable(summary(emm_ZIB_zero), digits = 3,
            caption = "Single-step ZIB (zi component): four-condition marginal means (Pr(y=0), reference cell = 3D-only·simple)") %>%
        kableExtra::kable_styling(full_width = FALSE)
    )

    kbl_show(
      knitr::kable(summary(pairs(emm_ZIB_zero)), digits = 3,
            caption = "Single-step ZIB (zi component): pairwise comparison of four conditions (Pr(y=0))") %>%
        kableExtra::kable_styling(full_width = FALSE)
    )
  } else {
    cat("The current version of emmeans does not support component = 'zi' and only reports point estimates for Pr(y=0).\n")
  }

} else {
  cat("ZIB is unavailable or has not converged, skipping this part.\n")
}
```



## ZIB Single Step: Strategy Difference Raincloud plot(E[y])

```{r ZIB_raincloud, fig.height=7, fig.width=7}
if(is.null(ZIB_best)){
  cat("The ZIB single-step model is not available, so the rainfall cloud map cannot be drawn.\n")
} else {

  # Marginal mean of E[y] (from outZ in the previous section)
  emm_ct_df <- outZ %>%
    dplyr::mutate(
      group = factor(sub("·.*", "", cell), levels = c("3D-only","2D+3D")),
      complexity = factor(sub(".*·", "", cell), levels = c("simple","complex"))
    ) %>%
    dplyr::transmute(
      group, complexity,
      time_sec  = `E[y]`,
      lower_sec = lo,
      upper_sec = hi
    )

  # Trial-level raw data: strategy variability
  dat_ct_raw <- df %>%
    dplyr::select(participant, group, complexity, strategy_div) %>%
    dplyr::rename(time_sec = strategy_div)


# ---------- color ----------
col_blue <- "#3E69B0"  # 3D-only
col_red  <- "#C61B21"  # 2D+3D

# ---------- Data range ----------
y_raw_min <- min(dat_ct_raw$time_sec, emm_ct_df$lower_sec, na.rm = TRUE)
y_raw_max <- max(dat_ct_raw$time_sec, emm_ct_df$upper_sec, na.rm = TRUE)

y_breaks_all  <- pretty(c(y_raw_min, y_raw_max), n = 5)
y_breaks_used <- y_breaks_all

y_axis_min <- min(y_breaks_used)
y_axis_max <- max(y_breaks_used)
dy         <- y_axis_max - y_axis_min

y_plot_min <- y_axis_min - 0.30 * dy
y_plot_max <- y_axis_max + 0.05 * dy

x_simple  <- 1
x_complex <- 2
x_rng     <- x_complex - x_simple

x_plot_min <- x_simple - 0.7
x_plot_max <- x_complex + 0.5

y_axis_x  <- x_simple - 0.35      
y_tick_x1 <- y_axis_x               
y_tick_x2 <- y_axis_x - 0.05        

x_axis_y  <- y_axis_min - 0.08 * dy      
x_tick_y1 <- x_axis_y                   
x_tick_y2 <- x_axis_y - 0.04 * dy         

# =====================================================================
# raincloud plot
# =====================================================================
p_ct_raincloud <- ggplot() +

  # ===== simple  =====
  gghalves::geom_half_violin(
    data = dat_ct_raw %>% filter(complexity == "simple", group == "3D-only"),
    aes(x = as.numeric(complexity), y = time_sec,
        fill = group, colour = group),
    side = "l", width = 0.4, trim = TRUE,
    alpha = 0.35, linewidth = 0.9
  ) +
  gghalves::geom_half_violin(
    data = dat_ct_raw %>% filter(complexity == "simple", group == "2D+3D"),
    aes(x = as.numeric(complexity), y = time_sec,
        fill = group, colour = group),
    side = "l", width = 0.4, trim = TRUE,
    alpha = 0.35, linewidth = 0.9
  ) +
  geom_point(
    data = dat_ct_raw %>% filter(complexity == "simple"),
    aes(x = as.numeric(complexity) + 0.10, y = time_sec, colour = group),
    position = position_jitter(width = 0.05, height = 0, seed = 2025),
    alpha = 0.6, shape = 16, size = 2.2
  ) +
  geom_boxplot(
    data = dat_ct_raw %>% filter(complexity == "simple", group == "3D-only"),
    aes(x = as.numeric(complexity) + 0.25, y = time_sec,
        fill = group, colour = group),
    width = 0.08,
    alpha = 0.25, outlier.shape = NA, linewidth = 0.9
  ) +
  geom_boxplot(
    data = dat_ct_raw %>% filter(complexity == "simple", group == "2D+3D"),
    aes(x = as.numeric(complexity) + 0.35, y = time_sec,
        fill = group, colour = group),
    width = 0.08,
    alpha = 0.25, outlier.shape = NA, linewidth = 0.9
  ) +

  # ===== complex  =====
  gghalves::geom_half_violin(
    data = dat_ct_raw %>% filter(complexity == "complex", group == "3D-only"),
    aes(x = as.numeric(complexity), y = time_sec,
        fill = group, colour = group),
    side = "r", width = 0.4, trim = TRUE,
    alpha = 0.35, linewidth = 0.9
  ) +
  gghalves::geom_half_violin(
    data = dat_ct_raw %>% filter(complexity == "complex", group == "2D+3D"),
    aes(x = as.numeric(complexity), y = time_sec,
        fill = group, colour = group),
    side = "r", width = 0.4, trim = TRUE,
    alpha = 0.35, linewidth = 0.9
  ) +
  geom_point(
    data = dat_ct_raw %>% filter(complexity == "complex"),
    aes(x = as.numeric(complexity) - 0.10, y = time_sec, colour = group),
    position = position_jitter(width = 0.05, height = 0, seed = 2025),
    alpha = 0.6, shape = 16, size = 2.2
  ) +
  geom_boxplot(
    data = dat_ct_raw %>% filter(complexity == "complex", group == "3D-only"),
    aes(x = as.numeric(complexity) - 0.35, y = time_sec,
        fill = group, colour = group),
    width = 0.08,
    alpha = 0.25, outlier.shape = NA, linewidth = 0.9
  ) +
  geom_boxplot(
    data = dat_ct_raw %>% filter(complexity == "complex", group == "2D+3D"),
    aes(x = as.numeric(complexity) - 0.25, y = time_sec,
        fill = group, colour = group),
    width = 0.08,
    alpha = 0.25, outlier.shape = NA, linewidth = 0.9
  ) +

  # ===== EMM =====
  geom_line(
  data = emm_ct_df,
  aes(
    x = as.numeric(complexity) + case_when(
      group == "3D-only" & complexity == "simple" ~ 0.25,  
      group == "2D+3D" & complexity == "simple" ~ 0.35,  
      group == "3D-only" & complexity == "complex" ~ -0.35, 
      group == "2D+3D" & complexity == "complex" ~ -0.25  
    ),
    y = time_sec,
    group = group,
    colour = group
  ),
  linewidth = 1.1
) +
geom_errorbar(
  data = emm_ct_df,
  aes(
    x = as.numeric(complexity) + case_when(
      group == "3D-only" & complexity == "simple" ~ 0.25,
      group == "2D+3D" & complexity == "simple" ~ 0.35,
      group == "3D-only" & complexity == "complex" ~ -0.35,
      group == "2D+3D" & complexity == "complex" ~ -0.25
    ),
    ymin = lower_sec, ymax = upper_sec,
    group = group, colour = group
  ),
  width = 0, linewidth = 1.1
) +
geom_point(
  data = emm_ct_df,
  aes(
    x = as.numeric(complexity) + case_when(
      group == "3D-only" & complexity == "simple" ~ 0.25,
      group == "2D+3D" & complexity == "simple" ~ 0.35,
      group == "3D-only" & complexity == "complex" ~ -0.35,
      group == "2D+3D" & complexity == "complex" ~ -0.25
    ),
    y = time_sec,
    group = group,
    fill = group,
    colour = group
  ),
  shape = 23, size = 3.0, stroke = 1.0
) +
  stat_summary(
    data = dat_ct_raw,
    aes(
      x = as.numeric(complexity) + dplyr::case_when(
        group == "3D-only" & complexity == "simple"  ~  0.25,
        group == "2D+3D"  & complexity == "simple"  ~  0.35,
        group == "3D-only" & complexity == "complex" ~ -0.35,
        group == "2D+3D"  & complexity == "complex" ~ -0.25
      ),
      y     = time_sec,
      colour = group,
      group  = interaction(group, complexity) 
    ),
    fun.min = min,      
    fun.max = max,      
    geom    = "errorbar",
    width   = 0.07,    
    linewidth = 0.9
  )+

  # =================================================================
  # Custom Axis
  # =================================================================
  geom_segment(
    inherit.aes = FALSE,
    aes(x = y_axis_x, xend = y_axis_x,
        y = y_axis_min, yend = y_axis_max),
    linewidth = 0.7
  ) +
  geom_segment(
    inherit.aes = FALSE,
    data = data.frame(y = y_breaks_used),
    aes(x = y_tick_x1, xend = y_tick_x2,
        y = y, yend = y),
    linewidth = 0.7
  ) +


  geom_segment(
    inherit.aes = FALSE,
    aes(x = x_simple, xend = x_complex,
        y = x_axis_y, yend = x_axis_y),
    linewidth = 0.7
  ) +
  geom_segment(
    inherit.aes = FALSE,
    data = data.frame(x = c(x_simple, x_complex)),
    aes(x = x, xend = x,
        y = x_tick_y1, yend = x_tick_y2),
    linewidth = 0.7
  ) +

  # =================================================================
  # Manually draw scale numbers and axis titles near the new axis
  # =================================================================

  geom_text(
    inherit.aes = FALSE,
    data = data.frame(
      y     = y_breaks_used,
      label = y_breaks_used
    ),
    aes(x = y_axis_x - 0.15, y = y, label = label),
    hjust = 0, vjust = 0.5, size = 3.5
  ) +

  geom_text(
    inherit.aes = FALSE,
    data = data.frame(
      x     = c(x_simple, x_complex),
      label = c("Simple", "Complex")
    ),
    aes(x = x, y = x_tick_y2 - 0.03 * dy, label = label),
    vjust = 1, size = 3.5
  ) +

  geom_text(
    inherit.aes = FALSE,
    aes(x = y_axis_x - 0.3,
        y = (y_axis_min + y_axis_max) / 2,
        label = "Strategy divergence"),
    angle = 90, hjust = 0.5, vjust = 0.5, size = 4
  ) +

  geom_text(
    inherit.aes = FALSE,
    aes(x = (x_simple + x_complex) / 2,
        y = x_tick_y2 - 0.10 * dy,
        label = "Complexity"),
    vjust = 1, size = 4
  ) +

  scale_colour_manual(
    values = c("3D-only" = col_blue,
               "2D+3D"   = col_red),
    name   = "Condition"
  ) +
  scale_fill_manual(
    values = c("3D-only" = alpha(col_blue, 0.35),
               "2D+3D"   = alpha(col_red,  0.35)),
    name   = "Condition"
  ) +

  scale_y_continuous(
    limits = c(y_plot_min, y_plot_max),
    breaks = y_breaks_used
  ) +
  scale_x_continuous(
    limits = c(x_plot_min, x_plot_max),
    breaks = c(x_simple, x_complex),
    labels = NULL   # 不用默认标签，自己画了
  ) +

  # ===== theme =====
  theme_minimal(base_size = 12) +
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.major.y = element_blank(),   
    panel.grid.minor   = element_blank(),
    axis.text.x        = element_blank(),   
    axis.text.y        = element_blank(),
    axis.title.x       = element_blank(),  
    axis.title.y       = element_blank(),
    axis.ticks         = element_blank(),
    axis.line          = element_blank(),
    legend.position    = "top",
    legend.title       = element_text(size = 11),
    legend.margin      = margin(b = -5),
    plot.margin        = margin(10, 25, 35, 45)
  ) +
  coord_cartesian(clip = "off")

p_ct_raincloud

}
```

# Two-step synthesis estimates

```{r}
emm_zero <- emmeans(L_best, ~ group*complexity, type = "response")
emm_pos  <- emmeans(P_best, ~ group*complexity, type = "response")

# ——Overall effect (Wald test, similar to factor ANOVA)——
joint_zero <- joint_tests(emm_zero)
joint_pos  <- joint_tests(emm_pos)

kbl_show(
  knitr::kable(as.data.frame(joint_zero), digits = 3,
        caption = "Zero/Non-zero levels: Overall Wald test for group/complexity/interaction") %>%
    kableExtra::kable_styling(full_width = FALSE)
)

kbl_show(
  knitr::kable(as.data.frame(joint_pos), digits = 3,
        caption = "Level: Overall Wald Test for group / complexity / interaction") %>%
    kableExtra::kable_styling(full_width = FALSE)
)

# ——Estimates for four conditions  Pairwise comparisons between conditions——
kbl_show(
  knitr::kable(summary(emm_zero), digits = 3,
        caption = "Zero/Non-zero layers: Marginal means of the four conditions (probability scale)") %>%
    kableExtra::kable_styling(full_width = FALSE)
)

kbl_show(
  knitr::kable(summary(pairs(emm_zero)), digits = 3,
        caption = "Zero/Non-zero Layers: Pairwise Comparison of Four Conditions (Probability Scale)") %>%
    kableExtra::kable_styling(full_width = FALSE)
)

kbl_show(
  knitr::kable(summary(emm_pos), digits = 3,
        caption = paste0("Positive layer (", pos_family, "): the marginal mean of the four conditions (expected value scale)")) %>%
    kableExtra::kable_styling(full_width = FALSE)
)

kbl_show(
  knitr::kable(summary(pairs(emm_pos)), digits = 3,
        caption = paste0("Positive Layer (", pos_family, "): Comparison of four conditions (expected value scale)")) %>%
    kableExtra::kable_styling(full_width = FALSE)
)
```

```{r}
# ——Parameter sampling to synthesize the overall expected value E[y]——
cells2 <- expand.grid(group = c("3D-only","2D+3D"),
                      complexity = c("simple","complex")) %>%
  mutate(
    group = factor(group, levels = c("3D-only","2D+3D")),
    complexity = factor(complexity, levels = c("simple","complex")),
    cell = paste(group, complexity, sep = "·")
  )

X <- model.matrix(~ group*complexity, data = cells2)

fix_L <- fixef(L_best)$cond
V_L   <- as.matrix(vcov(L_best)$cond)

fix_P <- fixef(P_best)$cond
V_P   <- as.matrix(vcov(P_best)$cond)

inv_logit <- function(eta) 1/(1+exp(-eta))

B <- params$B
set.seed(2025)
bt_L <- MASS::mvrnorm(B, mu = fix_L, Sigma = V_L)
bt_P <- MASS::mvrnorm(B, mu = fix_P, Sigma = V_P)

p_zero <- inv_logit(bt_L %*% t(X))
mu_pos <- if(pos_family == "Beta") inv_logit(bt_P %*% t(X)) else exp(bt_P %*% t(X))
Ey <- (1 - p_zero) * mu_pos

summ <- function(M, metric_name, cell_vec){
  tibble(
    metric = metric_name,
    cell = cell_vec,
    mean = apply(M, 2, mean),
    lo = apply(M, 2, quantile, probs = .025),
    hi = apply(M, 2, quantile, probs = .975)
  )
}

out_zero <- summ(p_zero, "Pr(y=0)", cells2$cell)
out_pos  <- summ(mu_pos, "E[y|y>0]", cells2$cell)
out_all  <- summ(Ey, "E[y]", cells2$cell)

res <- bind_rows(out_zero, out_pos, out_all) %>%
  relocate(metric, cell, mean, lo, hi)

kbl_show(
  knitr::kable(res, digits = 3,
        caption = paste0("Two-step synthesis: four-condition estimate and 95%（B = ", B, ")")) %>%
    kableExtra::kable_styling(full_width = FALSE)
)
```

# Comparison: ZIB vs Two-Step
```{r}
plot_df <- res %>%
  filter(metric == "E[y]") %>%
  mutate(source = "Two-step")

if(!is.null(ZIB_best)){
  pr0 <- predict(ZIB_best, newdata = cells, type = "zprob", re.form = NA)
  mu  <- predict(ZIB_best, newdata = cells, type = "conditional", re.form = NA)
  plot_df <- bind_rows(
    plot_df,
    tibble(cell = cells$cell,
           mean = (1 - pr0) * mu,
           lo = NA_real_, hi = NA_real_,
           source = "ZIB")
  )
}

ggplot(plot_df, aes(x = cell, y = mean, color = source, group = source)) +
  geom_point(position = position_dodge(width = .3), size = 2) +
  geom_line(position = position_dodge(width = .3)) +
  theme_minimal() +
  labs(x = NULL, y = "E[y]",
       title = "Overall expected E[y]: ZIB single-step vs two-step synthesis (control)")
```

# Summary
- **ZIB single-step** and **two-step** should be consistent in direction; in the report, choose one of the two to present as the main result;
- Three types of indicators: `Pr(zero)`, `E[y|y>0]`, `E[y]`, with sensitivity analysis possible in the appendix.

## 7. Strategy Divergence — Δ EMMs Forest (overall; integrated table + plot, aligned version)

> Use the overall metric m = (1 - p0) * mu_plus (0–1), with the horizontal axis as Δ(B − A) and the reference line = 0; the plotting framework is exactly the same as CT/HM/Difficulty.

```{r strategy_forest_overall_aligned, fig.width=10, fig.height=5, out.width='100%'}
suppressPackageStartupMessages({
  library(dplyr); library(tidyr); library(forcats)
  library(ggplot2); library(scales); library(glue); library(tibble); library(purrr)
})

# ===== Strategy Divergence (overall): m = (1 - p0) * mu_plus on 0–1 scale; report Δ (B - A), ref = 0 =====
# Assumptions:
# - ZIB_best exists earlier in the document.
# - Factors match other DVs: group = {3D-only, 2D+3D}, complexity = {simple, complex}.

cells <- tidyr::expand_grid(
  group = factor(c("3D-only","2D+3D"), levels = c("3D-only","2D+3D")),
  complexity = factor(c("simple","complex"), levels = c("simple","complex"))
) %>%
  dplyr::mutate(cell = paste(group, complexity, sep = "·"))

stopifnot(exists("ZIB_best"))

# 1) E[Y] and its SE via delta method (ignoring covariance between parts; conservative)
pr0 <- predict(ZIB_best, newdata = cells, type = "zprob",       se.fit = TRUE, re.form = NA)
mu  <- predict(ZIB_best, newdata = cells, type = "conditional", se.fit = TRUE, re.form = NA)

m    <- (1 - pr0$fit) * mu$fit                        # overall expectation, 0–1
se_m <- sqrt(((1 - pr0$fit)^2) * (mu$se.fit^2) + (mu$fit^2) * (pr0$se.fit^2))

# Directly add the m and se columns to the cells dataframe without separately referencing the column names.
m_tbl <- cells %>%
  dplyr::mutate(
    m = m,    # Overall Expectations
    se = se_m # Standard Error
  )

# 2) Four contrasts (B − A), keep same order/labels as other DVs
get_m  <- function(g, c) m_tbl$m [m_tbl$group==g & m_tbl$complexity==c]
get_se <- function(g, c) m_tbl$se[m_tbl$group==g & m_tbl$complexity==c]

contr_rows <- list(
  list(Contrast="Simple Scenario",        Block="Condition Effect (2D+3D vs. 3D-only)",
       A=list(g="3D-only", c="simple"), B=list(g="2D+3D", c="simple")),
  list(Contrast="Complex Scenario",       Block="Condition Effect (2D+3D vs. 3D-only)",
       A=list(g="3D-only", c="complex"), B=list(g="2D+3D", c="complex")),
  list(Contrast=" 3D-only Condition",  Block="Complexity Effect (Complex vs. Simple)",
       A=list(g="3D-only", c="simple"), B=list(g="3D-only", c="complex")),
  list(Contrast=" 2D+3D Condition",    Block="Complexity Effect (Complex vs. Simple)",
       A=list(g="2D+3D",   c="simple"), B=list(g="2D+3D",   c="complex"))
)

df_contr <- purrr::map_dfr(contr_rows, function(r){
  mA  <- get_m(r$A$g, r$A$c);  mB  <- get_m(r$B$g, r$B$c)
  seA <- get_se(r$A$g, r$A$c); seB <- get_se(r$B$g, r$B$c)
  delta <- mB - mA
  se    <- sqrt(seA^2 + seB^2)        # conservative (no covariance)
  z     <- delta / se
  p     <- 2 * pnorm(-abs(z))
  tibble(Contrast = r$Contrast, Block = r$Block,
         delta = delta, low = delta - 1.96*se, high = delta + 1.96*se,
         p_raw = p,
         # For left table A|B display: show B | A to match other DVs
         a_m = mB, b_m = mA)
}) %>%
  dplyr::mutate(p_holm = p.adjust(p_raw, method = "holm"))

# 3) Vertical layout (identical to CT/HM/Difficulty)
row_order <- c("Simple Scenario", "Complex Scenario",
               " 3D-only Condition", " 2D+3D Condition")

gap_within_block1 <- 0.20; gap_between_blocks <- 0.60; gap_within_block2 <- 0.20
y1 <- 0.3; y2 <- y1 - gap_within_block1; y3 <- y2 - gap_between_blocks; y4 <- y3 - gap_within_block2
row_y <- c("Simple Scenario"=y1,"Complex Scenario"=y2," 3D-only Condition"=y3," 2D+3D Condition"=y4)
row_map <- tibble(Contrast = factor(names(row_y), levels = row_order),
                  y = as.numeric(row_y[as.character(Contrast)]))

df_contr <- df_contr %>%
  dplyr::mutate(Contrast = factor(Contrast, levels = row_order)) %>%
  dplyr::left_join(row_map, by = "Contrast") %>%
  dplyr::arrange(dplyr::desc(y))

y_top    <- max(row_map$y); y_bottom <- min(row_map$y)

# 4) Left table columns (same positions), but labels in raw 0–1 scale
col_x <- c("Contrast" = 0.15, "EMMs A|B" = 1.45,
           "Mean Diff. [95% CI]" = 2.25, "p (Holm)" = 3.35)
xlim_tbl <- c(0.05, 3.6)

# Forest area region on same canvas
forest_gap_x <- 0.40; forest_width <- 2.20
x_forest_min <- xlim_tbl[2] + forest_gap_x
x_forest_max <- x_forest_min + forest_width

# 5) Forest x-range on 0–1 units (Δ), with reference line at 0
xr_raw <- range(c(df_contr$low, df_contr$high), na.rm = TRUE)
pad    <- max(diff(xr_raw) * 0.06, 0.02)
xlim_for <- c(xr_raw[1] - pad, xr_raw[2] + pad)

choose_step <- function(rng, target_ticks = 4){
  width <- diff(rng); cand <- c(0.02, 0.05, 0.10, 0.20)
  cand[which.min(abs(width/cand - target_ticks))]
}
step <- choose_step(xlim_for, 4)
low  <- floor(xlim_for[1]/step) * step
high <- ceiling(xlim_for[2]/step) * step
x_breaks_for <- seq(low, high, by = step)
if (xlim_for[1] <= 0 && xlim_for[2] >= 0) x_breaks_for <- sort(unique(c(x_breaks_for, 0)))

xf <- function(x) scales::rescale(x, to = c(x_forest_min, x_forest_max), from = xlim_for)

df_contr <- df_contr %>%
  dplyr::mutate(x_low = xf(low), x_high = xf(high), x_mid = xf(delta))

x_ticks <- tibble(brk = x_breaks_for, x = xf(x_breaks_for))
tick_lab <- scales::label_number(accuracy = step)(x_ticks$brk)

# 6) Headers / thick line / section labels (unchanged)
header_text_y   <- y_top + 0.65
header_line_gap <- 0.08
header_line_y   <- header_text_y - header_line_gap

header_df <- tibble(x = unname(col_x), y = header_text_y, label = names(col_x))
sec_offset_top <- 0.20; sec_offset_bottom <- 0.20
sec_labels <- tibble(
  label = c("Condition Effect (2D+3D vs. 3D-only)", "Complexity Effect (Complex vs. Simple)"),
  y     = c(row_y["Simple Scenario"] + sec_offset_top,
            row_y[" 3D-only Condition"] + sec_offset_bottom),
  x     = unname(col_x["Contrast"])
)

body_dy <- 0.15
df_contr  <- df_contr  %>% dplyr::mutate(y = y + body_dy)
sec_labels <- sec_labels %>% dplyr::mutate(y = y + body_dy)

# 7) Bottom axis and dashed line at 0
axis_gap <- 0.15; axis_y <- y_bottom - axis_gap
tick_up <- 0.08; tick_down <- 0.06; tick_label_gap <- 0.28; tick_label_y <- axis_y - tick_label_gap

ylims_lower <- axis_y - (tick_label_gap + tick_down + 0.18)
ylims_upper <- y_top + 1.90
ylims <- c(ylims_lower, ylims_upper)

vline_df <- if (xlim_for[1] <= 0 && xlim_for[2] >= 0) tibble(x = xf(0)) else tibble(x = numeric(0))

# 8) Left table values (three decimals for 0–1 units)
tbl_show <- df_contr %>%
  dplyr::transmute(
    Contrast, y,
    `EMMs A|B` = sprintf("%s | %s",
                         format(round(a_m, 3), nsmall = 3),
                         format(round(b_m, 3), nsmall = 3)),
    `Mean Diff. [95% CI]` = sprintf("%.3f [%.3f, %.3f]", delta, low, high),
    `p (Holm)` = ifelse(p_holm < .001, "< .001", format(round(p_holm, 3), nsmall = 3))
  ) %>%
  tidyr::pivot_longer(cols = c("Contrast","EMMs A|B","Mean Diff. [95% CI]","p (Holm)"),
                      names_to = "col", values_to = "val") %>%
  dplyr::mutate(x = unname(col_x[col]), is_bold = (col == "Contrast"))

# 9) Single-canvas draw (layout identical to CT/HM/Difficulty)
ggplot() +
  geom_text(data = sec_labels, aes(x = x, y = y, label = label),
            hjust = 0, vjust = 0, fontface = "bold", size = 4) +
  geom_text(data = tbl_show,
            aes(x = x, y = y, label = val, fontface = ifelse(is_bold,"bold","plain"), colour = "black"),
            hjust = 0, vjust = 0.5, size = 3.6, lineheight = 1.0) +
  geom_text(data = header_df, aes(x = x, y = y, label = label),
            hjust = 0, vjust = 0, size = 3.7, fontface = "bold") +
  scale_colour_identity() +
  geom_segment(aes(x = -Inf, xend = Inf, y = header_line_y, yend = header_line_y), linewidth = 1.0) +
  geom_segment(data = df_contr, aes(x = x_low, xend = x_high, y = y, yend = y), linewidth = 0.9) +
  geom_point(data = df_contr, aes(x = x_mid, y = y), size = 2.8) +
  geom_segment(data = vline_df, aes(x = x, xend = x), y = axis_y, yend = header_line_y,
               linetype = "dashed", linewidth = 0.6) +
  annotate("text", x = x_forest_min, y = header_text_y, label = "Mean Diff. [95% CI]",
           hjust = -0.8, vjust = 0, size = 3.7, fontface = "bold") +
  geom_segment(data = x_ticks, aes(x = x, xend = x, y = axis_y - tick_down, yend = axis_y + tick_up),
               linewidth = 0.6, lineend = "butt") +
  geom_text(data = x_ticks, aes(x = x, y = tick_label_y, label = tick_lab), vjust = 1, size = 3.2) +
  annotate("segment", x = x_forest_min-0.1, xend = x_forest_max, y = axis_y, yend = axis_y, linewidth = 1.0, lineend = "square") +
  coord_cartesian(xlim = c(xlim_tbl[1], x_forest_max + 0.05), ylim = c(ylims[1], ylims[2]), clip = "on") +
  theme_minimal(base_size = 12) +
  theme(axis.title = element_blank(), axis.text = element_blank(),
        panel.grid = element_blank(), plot.margin = margin(t = 6, r = 6, b = 6, l = 4))
```
