---
title: "Completion Time : Model Selection"
author: "CongLiu"
output:
  html_document:
    toc: true
    toc_float: true
    number_sections: true
    df_print: paged
    word_document: default
    pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  fig.width = 7,
  fig.height = 5,
  fig.align = "center",
  dpi = 300
)

suppressPackageStartupMessages({
library(tidyverse)
library(lme4)
library(lmerTest)
library(nlme)
library(performance)
})
```

## 1. Data import and preprocessing
```{r data}
# ---- Data location ----
# The raw dataset is stored outside this analysis folder (e.g., in a sibling directory).
# If your folder layout differs, update `candidate_dirs` below.
data_file <- "Experimental_Raw_Data.csv"
candidate_dirs <- c(
  file.path("..", "data"),
  file.path("..", "datasets"),
  ".."
)
candidate_paths <- file.path(candidate_dirs, data_file)
data_path <- candidate_paths[file.exists(candidate_paths)][1]

if (is.na(data_path)) {
  stop("Data file not found. Checked:
  - ", paste(candidate_paths, collapse = "
  - "))
}

dat <- readr::read_csv(data_path, show_col_types = FALSE) %>% 
  mutate(
    group       = factor(group, levels = c("3D-only","2D+3D")),
    complexity  = factor(complexity, levels = c("simple","complex")),
    participant = factor(participant),
    log_ct      = log(completion_time)
  )
```

## 2. Candidate model specification and fitting
| Model ID | Random-effects structure | Heteroscedasticity | Implementation |
|---------|----------|--------|------|
| **M0**  | (1 &#124; participant) | No | `lmer` |
| **M1**  | (1 + complexity &#124; participant) | No | `lmer` |
| **M2**  | (1 &#124; participant) | varIdent(~1 &#124; complexity) | `nlme::lme` |
| **M3**  | (1 + complexity &#124; participant) | varIdent(~1 &#124; complexity) | `nlme::lme` |

```{r models}
# M0: random intercept
m0 <- lmer(log_ct ~ group * complexity + (1 | participant), data = dat, REML = TRUE)

# M1: random slope
m1 <- lmer(log_ct ~ group * complexity + (1 + complexity | participant), data = dat, REML = TRUE)

# M2: heteroscedastic (by complexity)
m2 <- lme(log_ct ~ group * complexity,
          random  = ~1 | participant,
          weights = varIdent(form = ~1 | complexity),
          data = dat,
          method = "REML")

# M3: random slope + heteroscedastic
m3 <- lme(log_ct ~ group * complexity,
          random  = ~1 + complexity | participant,
          weights = varIdent(form = ~1 | complexity),
          data = dat,
          control = lmeControl(returnObject = TRUE),
          method = "REML")
```

## 3. Model comparison
```{r compare}
models <- list(M0 = m0, M1 = m1, M2 = m2, M3 = m3)
performance::compare_performance(models, rank = TRUE)
```

```{r aic}
AICtab <- data.frame(
  Model = names(models),
  AIC   = sapply(models, AIC),
  BIC   = sapply(models, BIC)
)
knitr::kable(AICtab, digits = 1)
```

## 4. Diagnostics (best model: M2)
```{r diag}
performance::check_model(m2)
```

> **Decision.** Based on `compare_performance` and AIC/BIC, **M2** achieved the best overall performance; therefore, all subsequent analyses use **M2** (random intercepts with a complexity-specific variance structure).
